---
title: "Takeda Merged dataset"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Report 

This is a report for merged dataset.

```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>

# Import data to merge  

Merge data   
Filter data  
```{r, eval=F}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'

obj1= readRDS(paste0(dir,"rds/MT30269.raw.23.05.16.rds"))
obj2= readRDS(paste0(dir,"rds/MT30271.raw.23.05.16.rds"))
obj3= readRDS(paste0(dir,"rds/MT30273.raw.23.05.16.rds"))

# ## merge 
obj.srt = merge(obj1, y = c(obj2, obj3), add.cell.ids = c('MT30269','MT30271','MT30273'))
obj.srt = subset(obj.srt, sample %in% c("CTL","TAK","COM"))
obj.srt@meta.data[,c('orig.ident','sample')] %>% table() 
obj.srt <- subset(obj.srt, predicted_doublet == 'False' & nCount_RNA >= 500 & nCount_RNA < 20000 & percent.mt <= 25)

```

```{r, eval=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  ## Check ribosomal gene ratio and add ribosomal gene percentage
  ribosomal_genes <- rownames(obj.srt)[grepl('^RP', rownames(obj.srt))]
  obj.srt[["percent.ribo"]] <- PercentageFeatureSet(obj.srt, pattern = "^RP")
  
  ## Define variables in metadata to regress
  vars_to_regress <- c('percent.ribo')
  
  # Step 2: Scale and normalize data
  ## Regress out the uninteresting sources of variation in the data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(object = obj.srt, vars.to.regress = vars_to_regress, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(paste0(dir,'rds/MT_merged.3_conditions.23.09.15.rds'))
```


**pre-processing information** 

UMI cutoff : 500~ 20000.

mitochondrial cutoff : 25.

other cutoffs : NA.

doublet removal (scrublet) : Y.

regression: ribosomal genes.


# Import merged-processed data 

MT_merged.3_conditions rds   
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
obj.srt = readRDS(paste0(dir,'rds/MT_merged.3_conditions.23.09.15.rds'))
```

```{r}
obj.srt@meta.data$sample = factor(obj.srt@meta.data$sample, levels = c("CTL","TAK","COM"))
```


# UMAP 

## UMAP by sample
   
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'orig.ident')
```


```{r, fig.width=11, fig.height=4}
DimPlot(obj.srt, group.by = 'orig.ident', split.by = 'orig.ident', ncol = 3)
```

## UMAP by treatment    
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'sample')
```

```{r, fig.width=11, fig.height=4}
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 3)
```

## Distribution of number of cells 

```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample')
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment')

```


```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

# Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}
# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```

# Clusters with the resolution of 0.4   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```


```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}

# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 0.8   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}

# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 1   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```



**Find markers**   

resolution : 0.2, 0.4, 0.8, 1   
```{r, eval=FALSE, echo=F}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/merged/Takeda_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Marker genes 

## 0.2  

Download cluster marker file
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/merged/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.4  

Download cluster marker file
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/merged/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.8  

Download cluster marker file
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/merged/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 1  

Download cluster marker file
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/merged/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```



# Annotation of clusters  

## 0.2   
```{r}
library(DT)

resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.4   
```{r}
library(DT)

resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.8     
```{r}
library(DT)

resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 1   
```{r}
library(DT)

resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


Kindly complete the cluster annotation that you want to select and send it to Sung for update.  



# Composition of each cluster based on % of cells from each sample   

One cluster : consist of 3 samples (% of sample1, % of sample2, % of sample3)  

## res 0.2 
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```



```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100)%>% DT::datatable()
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```



## res 1 
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```

# Finalized cluster annotation   

Clustering resolution : 1   
Number of clusters : 34   
Number of annotation group : 5    

```{r}
# Create the data frame
data <- data.frame(
  cluster = c("cluster0", "cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "cluster6", "cluster7", "cluster8", "cluster9", "cluster10", "cluster11", "cluster12", "cluster13", "cluster14", "cluster15", "cluster16", "cluster17", "cluster18", "cluster19", "cluster20", "cluster21", "cluster22", "cluster23", "cluster24", "cluster25", "cluster26", "cluster27", "cluster28", "cluster29", "cluster30", "cluster31", "cluster32", "cluster33"),
  Proposed_cell_type = c(
    "mesenchymal", "lymphoid", "mesenchymal", "lymphoid", "lymphoid", "myeloid", "lymphoid", "mesenchymal", "lymphoid", "myeloid",
    "lymphoid", "lymphoid", "myeloid", "lymphoid", "lymphoid", "lymphoid", "lymphoid", "mesenchymal", "mesenchymal", "lymphoid",
    "myeloid", "lymphoid", "mesenchymal", "RBCs", "mesenchymal", "lymphoid", "low transcript", "mesenchymal", "low transcript",
    "lymphoid", "myeloid", "myeloid", "myeloid", "mesenchymal"
  )
)

# Print the data frame
data %>% DT::datatable(options = list(pageLength=length(unique(data$cluster))), rownames = F)
data %>% select(Proposed_cell_type) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("Group", "Number of clusters for the group"))
```


```{r}
obj.srt@meta.data$annotation = obj.srt@meta.data$RNA_snn_res.1 
levels(obj.srt@meta.data$annotation) = data$Proposed_cell_type
```


## UMAP   

```{r,fig.width=7.5, fig.height=6, out.width='80%'}
color= "annotation"
DimPlot(obj.srt, group.by = color)
```

```{r,fig.width=7.5, fig.height=6, out.width='80%'}
color= "annotation"
DimPlot(obj.srt, group.by = color, label = T, label.size = 3)
```


## Number of cells    


```{r}
color="annotation"
obj.srt@meta.data %>%
  group_by(!!sym(color)) %>%
  summarise(count = n(), .groups = "drop") %>% 
  DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data$annotation))))
```



```{r}
color="annotation"
obj.srt@meta.data %>%
  group_by(!!sym(color), orig.ident) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(!!sym(color), .add = TRUE) %>%
  DT::datatable(options = list(pageLength=15))
```


```{r, fig.width=12, fig.height=8}
obj.srt@meta.data %>%
  group_by(!!sym(color), orig.ident) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(!!sym(color)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(color), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r, fig.width=12, fig.height=8}
obj.srt@meta.data %>%
  group_by(!!sym(color), orig.ident) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(!!sym(color)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(color), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  labs(y = "Percentage") +
  theme_minimal()
```


# Next Step    

Generating new UMAPs for these 3 categories    

<br> 


lymphoid   
mesenchymal    
myeloid    



# Subset three cell types   
   
   

```{r}
celltypes_to_subset = c("lymphoid", "mesenchymal", "myeloid")
cat("celltypes_to_subset are ",celltypes_to_subset)
```

   
```{r, eval=FALSE}
celltypes_to_subset <- c("lymphoid", "mesenchymal", "myeloid")
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda/"

for (cell_type in celltypes_to_subset) {
  subset_obj <- subset(obj.srt, annotation == cell_type)
  rds_file <- file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", cell_type, ".rds"))
  saveRDS(subset_obj, rds_file)
}
```


Three subset objects have been saved in the designated folder.  


