---
title: "Takeda lymphoid from Merged dataset"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Report 

This is a report for lymphoid subset from merged dataset.

```{r, echo=F}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>

# Import data   


```{r}
celltypes_to_subset = c("lymphoid", "mesenchymal", "myeloid")
cat("Three cell types are ", "\n")
cat(celltypes_to_subset, "\n") 
cat("\n")
cat("The dataset here is ", celltypes_to_subset[1], "\n")

```


```{r, eval=FALSE}
# Preprocessing (happened only one time)
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda/"
obj.srt= readRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[1], ".rds")))
```

```{r, eval=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  ## Check ribosomal gene ratio and add ribosomal gene percentage
  ribosomal_genes <- rownames(obj.srt)[grepl('^RP', rownames(obj.srt))]
  obj.srt[["percent.ribo"]] <- PercentageFeatureSet(obj.srt, pattern = "^RP")
  
  ## Define variables in metadata to regress
  vars_to_regress <- c('percent.ribo')
  
  # Step 2: Scale and normalize data
  ## Regress out the uninteresting sources of variation in the data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(object = obj.srt, vars.to.regress = vars_to_regress, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[1], ".rds")))
```
<br>


**pre-processing information** 

UMI cutoff : 500~ 20000.

mitochondrial cutoff : 25.

other cutoffs : NA.

doublet removal (scrublet) : Y.

regression: ribosomal genes.


# Import processed data  

MT_merged.3_conditions_lymphoid      
```{r}
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda/"
obj.srt = readRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[1], ".rds")))
```

```{r}
obj.srt@meta.data$sample = factor(obj.srt@meta.data$sample, levels = c("CTL","TAK","COM"))
```


```{r}
# Create a data frame from the given data
data <- data.frame(
  Sample = c("30269", "30271","30273"),
  Condition = c("CTL,PEM,COM","CTL,PEM,COM","CTL,PEM,COM"),
  responding=c("NR","R","NR")
)

# Set the row names to the "RowNames" column
rownames(data) <- data$RowNames

# Print the data frame
data %>% gt::gt()

```
```{r}
obj.srt@meta.data$responding_group=factor(obj.srt@meta.data$orig.ident)
levels(obj.srt@meta.data$responding_group) = c("NR","R","NR")
```



# UMAP 

## UMAP by sample
   
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'orig.ident')
```


```{r,fig.width=14, fig.height=5}
DimPlot(obj.srt, group.by = 'orig.ident', split.by = 'orig.ident', ncol = 3)
```

## UMAP by treatment    
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'sample')
```

```{r,fig.width=14, fig.height=5}
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 3)
```

## UMAP by responding group
   
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'responding_group')
```


```{r,fig.width=11, fig.height=5}
DimPlot(obj.srt, group.by = 'responding_group', split.by = 'responding_group', ncol = 2)
```


```{r,fig.width=11, fig.height=5}
DimPlot(obj.srt, group.by = 'responding_group', split.by = 'orig.ident', ncol = 3)
```



## Distribution of number of cells 

```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample')
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment')

```


```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

# Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}
# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```

# Clusters with the resolution of 0.4   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```


```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}


obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 0.8   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}


obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 1   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```



**Find markers**   

resolution : 0.2, 0.4, 0.8, 1   
```{r, eval=FALSE, echo=F}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Marker genes 

## 0.2  

Download cluster marker file
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.4  

Download cluster marker file
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.8  

Download cluster marker file
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 1  

Download cluster marker file
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```



# Annotation of clusters  

## 0.2   
```{r}
library(DT)

resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.4   
```{r}
library(DT)

resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.8     
```{r}
library(DT)

resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 1   
```{r}
library(DT)

resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


Kindly complete the cluster annotation that you want to select and send it to Sung for update.  



# Composition of each cluster based on % of cells from each sample   

One cluster : consist of 3 samples (% of sample1, % of sample2, % of sample3)  

## res 0.2 
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```



```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100)%>% DT::datatable()
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```



## res 1 
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```


# Three major types of lymphocyte   
   
T cells, B cells, and natural killer cells   


```{r}
genes = c("CD3E","CD8A","CD4","CD79A", "NKG7")
FeaturePlot(obj.srt, features = genes, ncol = 3)
```


# Need of integration to annotate subtypes of lymphoid   

```{r, eval=FALSE}
# Use of Seurat v4 built-in integration method 
# Harmony (R version has the installation issue)
# Harmony (python version might be tried later)
# Tutorial URL : https://satijalab.org/seurat/articles/integration_introduction.html#perform-integration
obj.list <- SplitObject(obj.srt, split.by = "orig.ident")

# normalize and identify variable features for each dataset independently
ifnb.list <- lapply(X = obj.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = ifnb.list)

# Perform integration
immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, 
                                         anchor.features = features)
# this command creates an 'integrated' data assay
immune.combined <- IntegrateData(anchorset = immune.anchors)


# Perform an integrated analysis

# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(immune.combined) <- "integrated"

# Run the standard workflow for visualization and clustering
immune.combined <- ScaleData(immune.combined, verbose = FALSE)
immune.combined <- RunPCA(immune.combined, npcs = 30, verbose = FALSE)
immune.combined <- RunUMAP(immune.combined, reduction = "pca", dims = 1:30)
immune.combined <- FindNeighbors(immune.combined, reduction = "pca", dims = 1:30)
immune.combined <- FindClusters(immune.combined, resolution = c(0.1,0.2,0.5,0.8))

# Visualization
p1 <- DimPlot(immune.combined, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(immune.combined, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

DimPlot(immune.combined, group.by = "orig.ident", split.by = 'sample')
DimPlot(immune.combined, group.by = "integrated_snn_res.0.1", split.by = 'sample')
DimPlot(immune.combined, group.by = "orig.ident", split.by = 'sample')
DimPlot(immune.combined, group.by = "orig.ident", split.by = 'sample')


DefaultAssay(immune.combined) <- "RNA"
nk.markers <- FindConservedMarkers(immune.combined, ident.1 = 6, grouping.var = "orig.ident", verbose = FALSE)
head(nk.markers)

FeaturePlot(immune.combined, features = c("GNLY", "KLRC1"))

# Annotate cells and come back to the original dataset
```




# Next step   

Next steps will be here.. 







