---
title: "NK project: UMI 500~ 50000 MN1~MN4 Fibroblast"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---
```{r setup, include=FALSE, echo=T}
knitr::opts_chunk$set(echo=F, fig.align = "center", message=F, warning=F)
```
Report   
   
This is the initial analysis report of of Fibroblast cell analysis from MN1~MN4 merged dataset.

```{r libraries, echo=FALSE}
# Load required libraries
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```

* Import data  
Subseted UMI 500-50000, mitochondrial content cutoff 25%, singlet only     
 
```{r}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
## import data
obj.srt = readRDS(paste0(dir,('rds/MN1_MN4_subset.500.50000.25.singlet_Fibroblastsonly.23.10.23.rds')))
```



Snapshot of data   
```{r}
obj.srt@meta.data[1:3, 1:4]
```


# Overview of subset-data  

## Summary of data

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 50000 

mitochondrial cutoff : 25 

other cutoffs : NA 

doublet removal (scrublet) : Y 

Sample: MN1, MN2, MN3, MN4   



UMI and mitochondrial content   
```{r}
obj.srt@meta.data[,c('nCount_RNA','percent.mt')] %>% summary() 
```

Number of cells by sample   
```{r}
obj.srt@meta.data$orig.ident %>% table() %>% data.frame()
```

    
```{r, eval=FALSE}
cat("UMI and mitocondrial content % distribution ", "\n")
df= obj.srt@meta.data %>%
  group_by(orig.ident) %>%
  summarise(
    Mean_nCount_RNA = mean(nCount_RNA),
    Median_nCount_RNA = median(nCount_RNA),
    SD_nCount_RNA = sd(nCount_RNA),
    Q1_nCount_RNA = quantile(nCount_RNA, 0.25),  # 1st quantile
    Q3_nCount_RNA = quantile(nCount_RNA, 0.75),  # 3rd quantile
    Mean_percent_mt = mean(percent.mt),
    Median_percent_mt = median(percent.mt),
    SD_percent_mt = sd(percent.mt),
    Q1_percent_mt = quantile(percent.mt, 0.25),  # 1st quantile
    Q3_percent_mt = quantile(percent.mt, 0.75)  # 3rd quantile
  )
df %>% gt::gt()
```


```{r, eval=FALSE}
# Density plot of UMI counts  
obj.srt@meta.data %>% ggplot(aes(nCount_RNA, color=orig.ident)) + geom_density()
```


## Conditions of samples   


```{r}
# Create a data frame from the given data
data <- data.frame(
  Sample = c("MN1", "MN2","MN3", "MN4"),
  Cell_type = c("Fibroblasts", "Fibroblasts", "Fibroblasts", "Fibroblasts"),
  cGAMP = c("N", "Y", "N", "Y")
)

# Set the row names to the "RowNames" column
rownames(data) <- data$RowNames

# Print the data frame
data %>% gt::gt()

```


# UMAP  

## Samples on UMAP 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "orig.ident")
```


## Samples on UMAP in split mode  
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "orig.ident", split.by = "orig.ident",ncol = 2)
```


## Annotation on UMAP  

This annotation was done in merged dataset (as discussed).  


```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "annotation")
```


# Number of cells  
```{r}
obj.srt@meta.data[,"annotation"] %>% table() %>% data.frame() %>% gt::gt()
obj.srt@meta.data[,c("annotation","orig.ident")] %>% table() %>% data.frame() %>% tidyr::spread(orig.ident,Freq) %>% gt::gt()
```




```{r}
obj.srt@meta.data %>%
  group_by(orig.ident) %>%
  summarise(count = n()) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = orig.ident, y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal() +
  theme(legend.title = element_blank())

```


# Clusters on UMAP 

## Resolution of 0.2 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2", label = T, label.size = 3)
```


```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```



## Resolution of 0.4 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4", label = T, label.size = 3)
```


```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```



## Resolution of 0.8 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8", label = T, label.size = 3)
```


```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```



## Resolution of 1 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.1")
DimPlot(obj.srt, group.by = "RNA_snn_res.1", label = T, label.size = 3)
```

```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```


# Markers for clusters    

```{r, eval=FALSE}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)

# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/UMI500_50000/MN1_MN4_Fibroblast_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


```{r}
mks_set= "data/UMI500_50000/MN1_MN4_Fibroblast_"
```


## res 0.2  
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,mks_set, res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,mks_set, res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,mks_set, res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 1  
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}
mks = read.csv(paste0(dir,mks_set, res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```





<br>
<br>