---
title: "NK project: UMI 500~ 50000 MN5~MN6"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---
```{r setup, include=FALSE, echo=T}
knitr::opts_chunk$set(echo=F, fig.align = "center", message=F, warning=F)
```
Report   
   
This is the initial analysis report of of MN5~MN6 merged dataset.

```{r libraries, echo=FALSE}
# Load required libraries
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```

* Import data  
Subseted UMI 500-50000, mitochondrial content cutoff 25%, singlet only     
 
```{r}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
## import data
obj.srt = readRDS(paste0(dir,('rds/MN1_MN6_subset.500.50000.25.singlet.23.09.12.rds')))
```

* Subset data  

MN5,MN6 subsetting 
```{r, eval=FALSE}
# Happen only one time 
obj.srt = subset(obj.srt, orig.ident %in% paste0("MN", 5:6))

## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(paste0(dir,('rds/MN5_MN6_subset.500.50000.25.singlet.23.10.05.rds')))
```

# Import MN1~MN2 merged data  

```{r}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
## import data
obj.srt = readRDS(paste0(dir,('rds/MN5_MN6_subset.500.50000.25.singlet.23.10.05.rds')))
```

Snapshot of data   
```{r}
obj.srt@meta.data[1:3, 1:4]
```


# Overview of subset-data  

## Summary of data

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 50000 

mitochondrial cutoff : 25 

other cutoffs : NA 

doublet removal (scrublet) : Y 

Sample: MN3, MN4   



UMI and mitochondrial content   
```{r}
obj.srt@meta.data[,c('nCount_RNA','percent.mt')] %>% summary() 
```

Number of cells by sample   
```{r}
obj.srt@meta.data$orig.ident %>% table() %>% data.frame()
```

    
```{r, eval=FALSE}
cat("UMI and mitocondrial content % distribution ", "\n")
df= obj.srt@meta.data %>%
  group_by(orig.ident) %>%
  summarise(
    Mean_nCount_RNA = mean(nCount_RNA),
    Median_nCount_RNA = median(nCount_RNA),
    SD_nCount_RNA = sd(nCount_RNA),
    Q1_nCount_RNA = quantile(nCount_RNA, 0.25),  # 1st quantile
    Q3_nCount_RNA = quantile(nCount_RNA, 0.75),  # 3rd quantile
    Mean_percent_mt = mean(percent.mt),
    Median_percent_mt = median(percent.mt),
    SD_percent_mt = sd(percent.mt),
    Q1_percent_mt = quantile(percent.mt, 0.25),  # 1st quantile
    Q3_percent_mt = quantile(percent.mt, 0.75)  # 3rd quantile
  )
df %>% gt::gt()
```


```{r, eval=FALSE}
# Density plot of UMI counts  
obj.srt@meta.data %>% ggplot(aes(nCount_RNA, color=orig.ident)) + geom_density()
```


## Conditions of samples   


```{r}
# Create a data frame from the given data
data <- data.frame(
  Sample = c("MN5", "MN6"),
  HUVEC = c("HUVEC", "HUVEC"),
  Fibroblast = c("hLFB", "hLFB"),
  SCLC_line = c("H69", "H69"),
  cGAMP = c("N", "Y"),
  monocytes = c("N", "N"),
  NK_cell = c("Y", "Y")
)

# Set the row names to the "RowNames" column
rownames(data) <- data$RowNames

# Print the data frame
data %>% gt::gt()

```


# UMAP  

## Samples on UMAP 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "orig.ident")
```


## Samples on UMAP in split mode  
```{r, fig.width=12, fig.height=6}
DimPlot(obj.srt, group.by = "orig.ident", split.by = "orig.ident",ncol = 2)
```



## Clusters on UMAP 

### Resolution of 0.2 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2", label = T, label.size = 3)
```


```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```


```{r, eval=F}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```



### Resolution of 0.4 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4", label = T, label.size = 3)
```


```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```


```{r, eval=F}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```


### Resolution of 0.8 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8", label = T, label.size = 3)
```


```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```

```{r, eval=F}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```

### Resolution of 1 
```{r, fig.width=10, fig.height=8}
DimPlot(obj.srt, group.by = "RNA_snn_res.1")
DimPlot(obj.srt, group.by = "RNA_snn_res.1", label = T, label.size = 3)
```

```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res)] %>% table() %>% data.frame()
```


```{r, eval=F}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```


```{r, eval=FALSE}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)

# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/UMI500_50000/MN5_MN6_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Markers for clusters    


## res 0.2  
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN5_MN6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN5_MN6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN5_MN6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 1  
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN5_MN6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```



## Clsuter to cell type   
Please complete the table with your selections for cell types and associated genes for each cluster annotation. 
Following its completion, kindly send it to sungr_park@dfci.harvard.edu   


### resolution 0.2   
```{r}
resolution_number = 0.2 
res= paste0("RNA_snn_res.", resolution_number)

df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), cell_type = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


### resolution 0.4   
```{r}
resolution_number = 0.4 
res= paste0("RNA_snn_res.", resolution_number)

df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), cell_type = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


### resolution 0.8   
```{r}
resolution_number = 0.8 
res= paste0("RNA_snn_res.", resolution_number)

df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), cell_type = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


### resolution 1   
```{r}
resolution_number = 1 
res= paste0("RNA_snn_res.", resolution_number)

df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), cell_type = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```