---
title: "NK project: UMI 500~ 50000 MN3~MN4"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---
```{r setup, include=FALSE, echo=T}
knitr::opts_chunk$set(echo=F, fig.align = "center", message=F, warning=F)
```
Report   
   
This is the initial analysis report of of MN3~MN4 merged dataset.

```{r libraries, echo=FALSE}
# Load required libraries
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```

* Import data  
Subseted UMI 500-50000, mitochondrial content cutoff 25%, singlet only     
 
```{r, eval=FALSE}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
## import data
obj.srt = readRDS(paste0(dir,('rds/MN1_MN4_subset.500.50000.25.singlet.23.09.14.rds')))
```

* Subset data  

MN3,MN4 subsetting 
```{r, eval=FALSE}
# Happen only one time 
obj.srt = subset(obj.srt, orig.ident %in% paste0("MN", 3:4))

## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(paste0(dir,('rds/MN3_MN4_subset.500.50000.25.singlet.23.09.28.rds')))
```

# Import MN3~MN4 merged data  

```{r}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
## import data
obj.srt = readRDS(paste0(dir,('rds/MN3_MN4_subset.500.50000.25.singlet.23.09.28.rds')))
```

Snapshot of data   
```{r}
obj.srt@meta.data[1:3, 1:4]
```


# Overview of subset-data  

## Summary of data

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 50000 

mitochondrial cutoff : 25 

other cutoffs : NA 

doublet removal (scrublet) : Y 

Sample: MN3, MN4   



UMI and mitochondrial content   
```{r}
obj.srt@meta.data[,c('nCount_RNA','percent.mt')] %>% summary() 
```

Number of cells by sample   
```{r}
obj.srt@meta.data$orig.ident %>% table() %>% data.frame()
```

    
```{r, eval=FALSE}
cat("UMI and mitocondrial content % distribution ", "\n")
df= obj.srt@meta.data %>%
  group_by(orig.ident) %>%
  summarise(
    Mean_nCount_RNA = mean(nCount_RNA),
    Median_nCount_RNA = median(nCount_RNA),
    SD_nCount_RNA = sd(nCount_RNA),
    Q1_nCount_RNA = quantile(nCount_RNA, 0.25),  # 1st quantile
    Q3_nCount_RNA = quantile(nCount_RNA, 0.75),  # 3rd quantile
    Mean_percent_mt = mean(percent.mt),
    Median_percent_mt = median(percent.mt),
    SD_percent_mt = sd(percent.mt),
    Q1_percent_mt = quantile(percent.mt, 0.25),  # 1st quantile
    Q3_percent_mt = quantile(percent.mt, 0.75)  # 3rd quantile
  )
df %>% gt::gt()
```


```{r, eval=FALSE}
# Density plot of UMI counts  
obj.srt@meta.data %>% ggplot(aes(nCount_RNA, color=orig.ident)) + geom_density()
```


## Conditions of samples   


```{r}
# Create a data frame from the given data
data <- data.frame(
  Sample = c("MN3", "MN4"),
  HUVEC = c("HUVEC", "HUVEC"),
  Fibroblast = c("hLFB", "hLFB"),
  SCLC_line = c("CORL47", "CORL47"),
  cGAMP = c("N", "Y"),
  monocytes = c("Y", "Y"),
  NK_cell = c("Y", "Y")
)

# Set the row names to the "RowNames" column
rownames(data) <- data$RowNames

# Print the data frame
data %>% gt::gt()

```


# UMAP  

## Samples on UMAP 
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "orig.ident")
```


## Samples on UMAP in split mode  
```{r, fig.width=12, fig.height=6}
DimPlot(obj.srt, group.by = "orig.ident", split.by = "orig.ident",ncol = 2)
```


## Annotation on UMAP  

This annotation was done in merged dataset (as discussed).  


```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "annotation")
DimPlot(obj.srt, group.by = "annotation", label = T, label.size = 3)
```

```{r}
obj.srt@meta.data[,"annotation"] %>% table() %>% data.frame() %>% gt::gt()
obj.srt@meta.data[,c("annotation","orig.ident")] %>% table() %>% data.frame() %>% tidyr::spread(orig.ident,Freq) %>% gt::gt()
```



## Annotation on UMAP2   

```{r}
cell_types= unique(obj.srt@meta.data$annotation) 
samples = unique(obj.srt@meta.data$orig.ident)


color_sample_in_celltype = function(obj.srt, celltype){
  obj.srt@meta.data$color = "other"
  rows1= obj.srt@meta.data %>% filter(annotation == celltype & orig.ident==paste0("MN",3)) %>% rownames()
  rows2= obj.srt@meta.data %>% filter(annotation == celltype & orig.ident==paste0("MN",4)) %>% rownames()
  obj.srt@meta.data[rows1,]$color= paste0("MN",3)
  obj.srt@meta.data[rows2,]$color= paste0("MN",4)
  obj.srt@meta.data$color = factor(obj.srt@meta.data$color)
  
  obj.srt@meta.data$color %>% table()
  umap= obj.srt@reductions$umap@cell.embeddings %>% data.frame()
  p=umap %>% ggplot(aes(UMAP_1, UMAP_2, color=obj.srt@meta.data$color)) + geom_point(size=0.1) + 
    scale_color_manual(values = c("red","blue","#F2F4F4")) +theme_classic() + 
    theme(legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=3))) +ggtitle(celltype)
  print(p)
}

```


```{r,fig.width=12, fig.height=10}
color_sample_in_celltype(obj.srt, celltype = cell_types[1])
color_sample_in_celltype(obj.srt, celltype = cell_types[2])
color_sample_in_celltype(obj.srt, celltype = cell_types[3])
color_sample_in_celltype(obj.srt, celltype = cell_types[4])
color_sample_in_celltype(obj.srt, celltype = cell_types[5])

```



## Clusters on UMAP 

### Resolution of 0.2 
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.2", label = T, label.size = 3)
```

```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```



### Resolution of 0.4 
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.4", label = T, label.size = 3)
```

```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```


### Resolution of 0.8 
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8")
DimPlot(obj.srt, group.by = "RNA_snn_res.0.8", label = T, label.size = 3)
```


```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```

### Resolution of 1 
```{r, fig.width=12, fig.height=10}
DimPlot(obj.srt, group.by = "RNA_snn_res.1")
DimPlot(obj.srt, group.by = "RNA_snn_res.1", label = T, label.size = 3)
```

```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)

obj.srt@meta.data[,c(res,"annotation")] %>% table() %>% data.frame() %>% tidyr::spread(annotation, Freq) %>% DT::datatable(options = list(pageLength=length(unique(obj.srt@meta.data[,res]))), rownames = F)
```


```{r, eval=FALSE}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)

# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/UMI500_50000/MN3_MN4_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Markers for clusters    


## res 0.2  
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## res 1  
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
# res
```

Download cluster marker file
```{r}

mks = read.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```



# Characterization of clusters 

Please complete the table with your selections for cell types and associated genes for each cluster annotation. 
Following its completion, kindly send it to sungr_park@dfci.harvard.edu   

## res 0.2  
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), description = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len), rownames = F)
``` 


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), description = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len), rownames = F)
``` 


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), description = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len), rownames = F)
``` 


## res 1  
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
df =data.frame(cluster = paste0('Cluster ', levels(obj.srt@meta.data[,res])), description = rep("NA", length(levels(obj.srt@meta.data[,res]))), genes = rep("NA", length(levels(obj.srt@meta.data[,res]))))
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len), rownames = F)
``` 


# DEG between MN3 vs MN4 based on "ANNOTATION" from the merged dataset(MN1~MN4)  

```{r, eval=FALSE}
# Generate DEG files 
#as.vector(unique(obj.srt@meta.data$annotation)) %>% print()

## DEG gene analysis
mks.function= function(obj.srt, celltype='NK'){
  obj.srt@meta.data =obj.srt@meta.data %>% 
    mutate(compare=ifelse(annotation == celltype & orig.ident == 'MN3','g1',
                          ifelse(annotation == celltype & orig.ident == 'MN4','g2','other')))
  # add compare to the ident
  Idents(obj.srt) = 'compare'
  logfc=log2(1)
  mks =FindMarkers(obj.srt, ident.1 = 'g2', ident.2 = 'g1', 
                   logfc.threshold = logfc)
  pval=0.05
  fc=1.2
  mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                                 ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))
  mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))
  mks$gene = rownames(mks)
  mks =mks %>% mutate(labels= ifelse(DE == 'UP', gene, ifelse(DE=='DN',gene,'other')))
  mks =mks %>% arrange(desc(avg_log2FC))
  return(mks)
}

mks.nk=mks.function(obj.srt = obj.srt, celltype = 'NK cells')
mks.huvec=mks.function(obj.srt = obj.srt, celltype = 'HUVEC')
mks.fibro=mks.function(obj.srt = obj.srt, celltype = 'Fibroblasts')
mks.mono=mks.function(obj.srt = obj.srt, celltype = 'Monocytes')
mks.corl=mks.function(obj.srt = obj.srt, celltype = 'CORL47')

# save DEG files 
l <- list(NK = mks.nk, 
          HUVEC = mks.huvec, 
          Fibro = mks.fibro, 
          Mono = mks.mono, 
          CORL = mks.corl)

dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/data/UMI500_50000/'
lapply(names(l), function(name) {
  write.csv(l[[name]], paste0(dir, name, '.MN3.4.deg.csv'), row.names = T)
})

```

```{r}
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/data/UMI500_50000/'
list.files(dir, pattern = "deg.csv$") 
mks.nk= read.csv(paste0(dir,'NK','.MN3.4.deg.csv'), row.names = 1)
mks.huvec=read.csv(paste0(dir,'HUVEC','.MN3.4.deg.csv'), row.names = 1)
mks.fibro=read.csv(paste0(dir,'Fibro','.MN3.4.deg.csv'), row.names = 1)
mks.mono=read.csv(paste0(dir,'Mono','.MN3.4.deg.csv'), row.names = 1)
mks.corl=read.csv(paste0(dir,'CORL','.MN3.4.deg.csv'), row.names = 1)
```


## Download DEG files  

### NK DEGs  
```{r}
DT::datatable(mks.nk, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


### Fibroblast DEGs   
```{r}
DT::datatable(mks.fibro, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


### HUVEC DEGs     
```{r}
DT::datatable(mks.huvec, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


### Monocytes DEGs   
```{r}
DT::datatable(mks.mono, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


### CORL47 DEGs   
```{r}
DT::datatable(mks.corl, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```



## Volcanoplot   

```{r}
mks.volcanoplot = function(mks=mks, title){
  pval=0.05
  fc=1.2
  mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))
  mks =mks %>% mutate(labels= ifelse(!!sym("DE") == 'UP', gene, ifelse(!!sym("DE")=='DN',gene,'')))
  mks =mks %>% arrange(desc(avg_log2FC))
  p=mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color = DE)) + 
      geom_point(size = 1, alpha = 0.5) + 
      scale_color_manual(values = c("red",'blue', 'grey')) +
      theme_classic() +
      geom_vline(xintercept = c(-log2(fc), log2(fc)), color = 'grey') +
      geom_hline(yintercept = -log10(0.05), color = 'grey') +
      guides(colour = guide_legend(override.aes = list(size = 5))) +
      geom_text(aes(label = labels), size = 2, show.legend = FALSE, hjust = 0, nudge_x = 0.03) +
      ggeasy::easy_center_title() + ## to center title
      ggtitle(title)
  print(p)
}
```

```{r}
mks.volcanoplot.notext = function(mks=mks,title){
  pval=0.05
  fc=1.2
  mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))
  mks =mks %>% mutate(labels= ifelse(!!sym("DE") == 'UP', gene, ifelse(!!sym("DE")=='DN',gene,'')))
  mks =mks %>% arrange(desc(avg_log2FC))
  p=mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color = DE)) + 
      geom_point(size = 1, alpha = 0.5) + 
      scale_color_manual(values = c("red",'blue', 'grey')) +
      theme_classic() +
      geom_vline(xintercept = c(-log2(fc), log2(fc)), color = 'grey') +
      geom_hline(yintercept = -log10(0.05), color = 'grey') +
      guides(colour = guide_legend(override.aes = list(size = 5))) +
      ggeasy::easy_center_title() + ## to center title
      ggtitle(title)
  print(p)
}
```

```{r, fig.width=10, fig.height=6, out.width="120%"}
t= "NK"
mks.volcanoplot(mks.nk, title=t)
mks.volcanoplot.notext(mks.nk, title=t)
```


```{r, fig.width=10, fig.height=6, out.width="120%"}
t= "Fibroblast"
mks.volcanoplot(mks.fibro, title=t)
mks.volcanoplot.notext(mks.fibro, title=t)
```

```{r, fig.width=10, fig.height=6, out.width="120%"}
t= "HUVEC"
mks.volcanoplot(mks.huvec, title=t)
mks.volcanoplot.notext(mks.huvec, title=t)
```

```{r, fig.width=10, fig.height=6, out.width="120%"}
t= "Monocyte"
mks.volcanoplot(mks.mono, title=t)
mks.volcanoplot.notext(mks.mono,title=t)
```

```{r, fig.width=10, fig.height=6, out.width="120%"}
t= "CORL47"
mks.volcanoplot(mks.corl, title=t)
mks.volcanoplot.notext(mks.corl, title=t)
```


## Violin plot of UP/DN regulated genes 

Please provide your gene list for viloin plots.



# Common gene analysis: DEG between MN3 vs MN4 based on "ANNOTATION" from the merged dataset(MN1~MN4) 


## UP regulated genes  


```{r}
condition = "UP"
v1=mks.nk %>% filter(DE == condition) %>% select(gene) %>% pull()
v2=mks.fibro %>% filter(DE == condition) %>% select(gene) %>% pull()
v3=mks.huvec %>% filter(DE == condition) %>% select(gene) %>% pull()
v4=mks.mono %>% filter(DE == condition) %>% select(gene) %>% pull()
v5=mks.corl %>% filter(DE == condition) %>% select(gene) %>% pull()

ven_list = list(CORL47 = v5,
                HUVEC = v3,
                Fibroblast = v2,
                Monocyte = v4,
                NK = v1)

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```


```{r, fig.width=10, fig.height=8, out.width="70%"}
plot(ven_out)
```


```{r}
plot(ven_out, type = "upset")
```

**Download shared gene list**  
```{r}
DT::datatable(ven_out@result, editable = FALSE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


## DOWN regulated genes  


```{r}
condition = "DN"
v1=mks.nk %>% filter(DE == condition) %>% select(gene) %>% pull()
v2=mks.fibro %>% filter(DE == condition) %>% select(gene) %>% pull()
v3=mks.huvec %>% filter(DE == condition) %>% select(gene) %>% pull()
v4=mks.mono %>% filter(DE == condition) %>% select(gene) %>% pull()
v5=mks.corl %>% filter(DE == condition) %>% select(gene) %>% pull()

ven_list = list(CORL47 = v5,
                HUVEC = v3,
                Fibroblast = v2,
                Monocyte = v4,
                NK = v1)

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```


```{r, fig.width=10, fig.height=8, out.width="70%"}
plot(ven_out)
```


```{r}
plot(ven_out, type = "upset")
```


**Download shared gene list**  
```{r}
DT::datatable(ven_out@result, editable = FALSE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=20))
```


# GSEA   

```{r, eval=FALSE}
# Merge all DEG files. 
mks.nk$cell_type = "NK"
mks.fibro$cell_type = "Fibroblast"
mks.huvec$cell_type = "HUVEC"
mks.mono$cell_type = "Monocytes"
mks.corl$cell_type = "CORL47"

mks.all = rbind(mks.nk,
                mks.fibro,
                mks.huvec,
                mks.mono,
                mks.corl)
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
output_file = paste0(dir,"data/UMI500_50000/MN3_MN4_DEG_all.csv")
mks.all %>% write.csv(file = output_file, row.names = T)
```


```{r, eval=FALSE}
mks.all = read.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_DEG_all.csv"), row.names = 1)
```


```{r}
library(clusterProfiler)
hallmark <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% dplyr::select(gs_name, gene_symbol)
perform_GSEA <- function(res, ref, title, pvalueCutoff = 1) {
  ranking <- function(res) {
    df <- res$avg_log2FC
    names(df) <- res$gene
    df <- sort(df, decreasing = TRUE)
    return(df)
  }
  
  ranked.res <- ranking(res)
  
  x <- clusterProfiler::GSEA(geneList = ranked.res,
                             TERM2GENE = ref,
                             pvalueCutoff = pvalueCutoff,
                             pAdjustMethod = "BH",
                             verbose = TRUE,
                             seed = TRUE)
  
  result <- x@result %>% arrange(desc(NES))
  result <- result[, c('NES', 'pvalue', 'p.adjust', 'core_enrichment', 'ID')]
  result$celltype = title
  return(result)
}

```


```{r}
title='MONOCYTE'
gsea.mono = perform_GSEA(res = mks.mono, ref = hallmark, title = title)
title='HUVEC'
gsea.huvec = perform_GSEA(res = mks.huvec, ref = hallmark, title = title)
title='Fibroblast'
gsea.fibro = perform_GSEA(res = mks.fibro, ref = hallmark, title = title)
title='NK'
gsea.nk = perform_GSEA(res = mks.nk, ref = hallmark, title = title)
title='CORL47'
gsea.corl = perform_GSEA(res = mks.corl, ref = hallmark, title = title)
```


```{r, eval=FALSE}
## find the shared GSEA IDs
gsea.list = list(NK=gsea.nk$ID,
                 MONO=gsea.mono$ID,
                 HUVEC=gsea.huvec$ID,
                 CORL=gsea.corl$ID,
                 FIBRO=gsea.fibro$ID)
## generate venndiagram data
## generate venn diagram 
ven_gsea <- VennDetail::venndetail(gsea.list)
ven_gsea@result[1:3,]
IDs=ven_gsea@result[ven_gsea@result$Subset =='Shared',]$Detail

gsea.all=rbind(gsea.corl[IDs,c('ID','celltype','NES','pvalue','p.adjust')],
               gsea.huvec[IDs,c('ID','celltype','NES','pvalue','p.adjust')],
               gsea.mono[IDs,c('ID','celltype','NES','pvalue','p.adjust')],
               gsea.nk[IDs,c('ID','celltype','NES','pvalue','p.adjust')],
               gsea.fibro[IDs,c('ID','celltype','NES','pvalue','p.adjust')])
rownames(gsea.all) = paste0(gsea.all$ID,'_',gsea.all$celltype)
gsea.all %>% write.csv(paste0(dir,"data/UMI500_50000/MN3_MN4_GESA.HALLMARK.all.csv"))
```


## GSEA NES plot version 1   

```{r}
## GSEA plot by function
plot_gsea <- function(data, title) {
  p <- ggplot(data, aes(reorder(ID, NES), NES)) +
    geom_col(aes(fill=p.adjust)) +
    coord_flip() +
    labs(x="Pathway", y="Normalized Enrichment Score",
         title=paste0(title)) + 
    theme_classic() +
    scale_fill_gradient(low = 'red', high = '#E5E7E9') +
    theme(axis.text.x= element_text(size=5, face = 'bold'),
          axis.text.y= element_text(size=8, face = 'bold'), 
          axis.title =element_text(size=10))
  
  print(p)
}
```


```{r, fig.width=12, fig.height=8}
plot_gsea(gsea.nk, 'NK DEG HALLMARK pathways')
plot_gsea(gsea.huvec, 'HUVEC DEG HALLMARK pathways')
plot_gsea(gsea.fibro, 'FIBROBLASTS HALLMARK pathways')
plot_gsea(gsea.mono, 'MONOCYTES HALLMARK pathways')
plot_gsea(gsea.corl, 'CORL47 HALLMARK pathways')
```


## GSEA NES plot version 2   
```{r}
## GSEA plot version 2
plot_gsea2 <- function(data, title) {
  p <- ggplot(data, aes(reorder(ID, NES), NES)) +
    geom_col(aes(fill = ifelse(NES > 0, 'positively enriched', 'negatively enriched'))) +
    coord_flip() +
    labs(x = "Pathway", y = "Normalized Enrichment Score",
         title = paste0(title),
         fill = "Enrichment") +  # Add legend title here
    theme_classic() +
    scale_fill_manual(values = c('blue', 'red')) +
    theme(axis.text.x = element_text(size = 5, face = 'bold'),
          axis.text.y = element_text(size = 8, face = 'bold'), 
          axis.title = element_text(size = 10))
  
  print(p)
}
```

```{r, fig.width=12, fig.height=8}
plot_gsea2(gsea.nk, 'NK DEG HALLMARK pathways')
plot_gsea2(gsea.huvec, 'HUVEC DEG HALLMARK pathways')
plot_gsea2(gsea.fibro, 'FIBROBLASTS HALLMARK pathways')
plot_gsea2(gsea.mono, 'MONOCYTES HALLMARK pathways')
plot_gsea2(gsea.corl, 'CORL47 HALLMARK pathways')
```


## GSEA all   

```{r}
# read csv 
dir= '~/Desktop/DF/DFCI_Barbie/DFCI_Barbie_NK_Marco/'
file_path <- paste0(dir, "data/UMI500_50000/MN3_MN4_GESA.HALLMARK.all.csv")

gsea.all = read.csv(file_path, row.names = 1)
```

```{r}
custom_palette <- colorRampPalette(c("#2874A6", "white", "#FC4119"))(1000)

# descending order of NES
df =gsea.all %>% select(celltype,ID, NES) %>% tidyr::spread(ID, NES)
colsm = colSums(df[,2:ncol(df)])
colsm <- colsm[order(-colsm)]
#names(colsm)

# Get the order of the names in colsm
order_x <- names(colsm)

# Convert ID column to a factor with the desired order
gsea.all$ID <- factor(gsea.all$ID, levels = order_x)
```

```{r, fig.width=12, fig.height=7}
# Plot the ggplot with the reordered x-axis
gsea.all %>%
  ggplot(aes(x = ID, y = celltype)) +
  geom_point(aes(size = -(p.adjust), color = NES)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_color_gradientn(colors = custom_palette) +
  labs(color = "NES", size = "FDR")
```
