---
title: "2024 Daiichi DS1062a"
subtitle: "bulk RNA-seq"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(DT)
```

### Introduction

Paweletz lab  
2024.06.10
<br> 

[**2024_Daiichi_DS1062a_bulk**](Daiichi_2024_RNAseq_introduction.html)

* Experimental Conditions
* Raw data locations
* Dropbox link
* MultiQC report


## Analysis Report 


### Patient 30342  


* [**PCA & Data matrix (combined report)**](30342/Daiichi_2024_bulk_30342_PCA_TPM_combined.html)
* [**PCA Plot**](30342/Daiichi_2024_bulk_30342_PCA.html)
* [**TPM matrix/count table download**](30342/Daiichi_2024_bulk_30342_TPM.html)
* [**ssGSEA**](30342/Daiichi_2024_bulk_30342_ssGSEA_update.24.06.25.html)
  * HALLMARK pathways
  * scRNA-seq celltype marker genes
<!-- * [**Immune Cell Type Composition (TIMER test)**](30342/Daiichi_2024_bulk_30342_timer.html) -->
* [**DEG Analysis (1)**](30342/DEG/Daiichi_2024_bulk_30342_DEG1_update.24.06.25.html)
  * Dxd/DMSO
  * IgG-Dxd/DMSO
  * DS-1062a/DMSO<br>
  * [Comparison of 3 DEGs (UP/DN)](30342/DEG/Daiichi_2024_bulk_30342_DEG1_comparison_update.html)
  <br>
* [**DEG Analysis (2)**](30342/DEG/Daiichi_2024_bulk_30342_DEG2_DS1062a_comparions.html)
  * DS-1062a/Dxd
  * DS-1062a/IgG-Dxd
  * Comprison of 2 DEGs (UP/DN)<br>
* [**DEG Analysis (3)**](30342/DEG/Daiichi_2024_bulk_30342_DEG_3.html)
  * IgG-Dxd/Dxd

<!-- * [Cancer Cell feature scoring in samples (preparing)]() -->



### Patient 30348  

* [**PCA Plot**](30348/Daiichi_2024_bulk_30348_PCA.html)
* [**TPM matrix/count table download**](30348/Daiichi_2024_bulk_30348_TPM.html)
* [**ssGSEA**](30348/Daiichi_2024_bulk_30348_ssGSEA.html)
  * HALLMARK pathways
  * scRNA-seq celltype marker genes
<!-- * [**DEG Analysis (1)**](30348/DEG/Daiichi_2024_bulk_30348_DEG.html) -->
<!--   * Dxd/DMSO -->
<!--   * IgG-Dxd/DMSO -->
<!--   * DS-1062a/DMSO<br> -->
<!--   * [Comparison of 3 DEGs (UP/DN)](30348/DEG/Daiichi_2024_bulk_30348_DEG_DS1062a_comparions.html) -->
<!--   <br> -->




