---
title: "2024 Daiichi DS-1062a"
subtitle: "bulk RNA-seq"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: show 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(DT)
```


```{r}
dir = "~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_bulk_RNAseq/"
```


```{r}
count_TPM_matrix_all = readRDS(paste0(dir,"raw_data/30342/star30342_rawData.rds"))
```


```{r}
counts = count_TPM_matrix_all$count

# rename column names 
colnames(counts)= sub("_Count", "", colnames(counts))
rownames(counts) = counts$SYMBOL
counts = counts[,-1]

# Remove genes with no expression across samples 
counts = counts[rowSums(counts) !=0,]
```


```{r}
tpms = count_TPM_matrix_all$TPM

# rename column names 
colnames(tpms)= sub("_TPM", "", colnames(tpms))
rownames(tpms) = tpms$SYMBOL
tpms = tpms[,-1]

# match the rownames with counts 
tpms = tpms[rownames(counts),]
```

## DEG analysis

  * DS-1062a/Dxd
  * DS-1062a/IgG-Dxd

### DS-1062a vs Dxd {.tabset}

```{r}
library(DESeq2)
count.mtx = counts
cont = "Dxd"
tret = "DS1062a"
cols = c(c(4:6), c(10:12))
cond = c(rep(cont,3),
         rep(tret,3))
count.mtx = count.mtx[, cols]
# Generate info table
info <- data.frame(matrix(nrow = ncol(count.mtx), ncol = 2))
colnames(info) <- c('sample', 'cond')
info$sample <- colnames(count.mtx)
info$cond <- cond

# DESeq
dds <- DESeqDataSetFromMatrix(count.mtx, info, ~ cond)
dds <- DESeq(dds)
res <- results(dds)

res <- data.frame(res)
res4 <- res
```

#### Volcano plot (version1)
```{r}
fc = 1.5
pval = 0.05

add_DE_information <- function(res, fc = 1.5, pval = 0.05) {
  # Create DE_padj based on the padj column if it exists
  if ('padj' %in% colnames(res)) {
    res <- res %>% 
      mutate(DE_padj = ifelse(log2FoldChange >= log2(fc) & padj < pval, 'UP',
                              ifelse(log2FoldChange <= -log2(fc) & padj < pval, 'DN', 'no_sig')))
    
    res$DE_padj <- factor(res$DE_padj, levels = c('UP', 'DN', 'no_sig'))
  }
  
  # Create DE_pvalue based on the pvalue column if it exists
  if ('pvalue' %in% colnames(res)) {
    res <- res %>% 
      mutate(DE_pvalue = ifelse(log2FoldChange >= log2(fc) & pvalue < pval, 'UP',
                                ifelse(log2FoldChange <= -log2(fc) & pvalue < pval, 'DN', 'no_sig')))
    
    res$DE_pvalue <- factor(res$DE_pvalue, levels = c('UP', 'DN', 'no_sig'))
  }
  
  return(res)
}

res4 = add_DE_information(res = res4)

res4 %>% 
  ggplot(aes(log2FoldChange, -log10(padj))) + 
  geom_point(size=1, alpha=0.5) + 
  scale_color_manual(values = c("red3","royalblue3","grey"), guide = FALSE) +
  theme_classic() +
  geom_vline(xintercept = c(-log2(fc),log2(fc)), color='grey') +
  geom_hline(yintercept = -log10(pval),color='grey') 

```

#### Volcano plot (version2)
```{r}
plot_DE_information <- function(res, fc = 1.5, pval = 0.05, p_col = 'padj') {
  # Check if the provided p_col exists in the dataframe
  if (!(p_col %in% colnames(res))) {
    stop(paste("The column", p_col, "does not exist in the dataframe"))
  }
  
  # Create DE column based on the provided p_col
  if (p_col == 'padj') {
    res <- res %>% 
      mutate(DE_padj = ifelse(log2FoldChange >= log2(fc) & padj < pval, 'UP',
                              ifelse(log2FoldChange <= -log2(fc) & padj < pval, 'DN', 'no_sig')))
    res$DE_padj <- factor(res$DE_padj, levels = c('UP', 'DN', 'no_sig'))
    df <- res$DE_padj %>% table() %>% data.frame()
    up <- df[,2][1]
    dn <- df[,2][2]
    y_col <- 'padj'
    de_col <- 'DE_padj'
  } else if (p_col == 'pvalue') {
    res <- res %>% 
      mutate(DE_pvalue = ifelse(log2FoldChange >= log2(fc) & pvalue < pval, 'UP',
                                ifelse(log2FoldChange <= -log2(fc) & pvalue < pval, 'DN', 'no_sig')))
    res$DE_pvalue <- factor(res$DE_pvalue, levels = c('UP', 'DN', 'no_sig'))
    df <- res$DE_pvalue %>% table() %>% data.frame()
    up <- df[,2][1]
    dn <- df[,2][2]
    y_col <- 'pvalue'
    de_col <- 'DE_pvalue'
  } else {
    stop("Invalid p_col value. Use 'padj' or 'pvalue'.")
  }

  # Create the plot
  p <- res %>%
    ggplot(aes(log2FoldChange, -log10(get(y_col)), color=get(de_col))) + 
    geom_point(size=0.5, shape=19, alpha=0.7) +
    geom_vline(xintercept = c(-log2(fc), log2(fc)), size=0.1, color="grey88") +
    geom_hline(yintercept = -log10(pval), size=0.1) +
    scale_color_manual(values = c("red3","royalblue3","grey"), guide = FALSE) +
    theme_bw() +
    annotate("text", x = Inf, y = Inf, label = paste0("UP: ", up), 
             hjust = 1.1, vjust = 2, size = 5, color = "red") +
    annotate("text", x = -Inf, y = Inf, label = paste0("DN: ", dn), 
             hjust = -0.1, vjust = 2, size = 5, color = "royalblue") +
    theme_bw() +
    ylab(paste0("log10", "(",p_col, ")"))
  
  return(p)
}

plot_DE_information(res4, fc = 1.5, pval = 0.05, p_col = 'padj')
plot_DE_information(res4, fc = 1.5, pval = 0.05, p_col = 'pvalue')

```

#### Volcano plot (version3)
```{r}
p1 = plot_DE_information(res4, fc = 1.5, pval = 0.05, p_col = 'padj')

p2 = plot_DE_information(res4, fc = 1.5, pval = 0.05, p_col = 'pvalue')

p1 + scale_y_sqrt()
p2 + scale_y_sqrt()
```


#### DEG table
```{r}
res4 %>% DT::datatable(extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=10))
```

```{r}
library(clusterProfiler)
hallmark <- msigdbr::msigdbr(species = "Homo sapiens", category = "H") %>% 
  dplyr::select(gs_name, gene_symbol)

perform_GSEA <- function(res, ref, pvalueCutoff = 1) {
  ranking <- function(res) {
    df <- res$log2FoldChange
    names(df) <- rownames(res)
    df <- sort(df, decreasing = TRUE)
    return(df)
  }
  
  ranked.res <- ranking(res)
  set.seed(123)
  x <- clusterProfiler::GSEA(geneList = ranked.res,
                             TERM2GENE = ref,
                             pvalueCutoff = pvalueCutoff,
                             pAdjustMethod = "BH",
                             verbose = TRUE,
                             seed = TRUE)
  
  result <- x@result %>% arrange(desc(NES))
  result <- result[, c('NES', 'pvalue', 'p.adjust', 'core_enrichment', 'ID')]
  return(result)
}

# Modified GSEA NES plot 
gsea_nes_plot =function(gsea.res, title, fontsize.x = 5, fontsize.y = 6){
  gsea.res %>% ggplot(aes(reorder(ID, NES), NES)) +
    geom_col(aes(fill=sig), color="grey1", size=0.2) +
    coord_flip() +
    labs(x="Pathway", y="Normalized Enrichment Score",
         title= "GSEA") + 
    theme_classic() +
    # scale_fill_gradient(low = '#FF0000', high = '#E5E7E9') +
    scale_fill_manual(values = c("#FF0000","grey88")) +
    theme(axis.text.x= element_text(size=fontsize.x, face = 'bold'),
          axis.text.y= element_text(size=fontsize.y, face = 'bold'), 
          axis.title =element_text(size=10)) +ggtitle(title)
}

# Modified GSEA NES plot (p value color version) 
gsea_nes_plot_pval =function(gsea.res, title, fontsize.x = 5, fontsize.y = 6){
  gsea.res %>% ggplot(aes(reorder(ID, NES), NES)) +
  geom_col(aes(fill=-pvalue), color="grey1", size=0.2) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title= "GSEA") + 
  theme_classic() +
  # scale_fill_gradient(low = '#FF0000', high = '#E5E7E9') +
  scale_fill_gradient(low = "grey88",high ="#FF0000" ) +
  theme(axis.text.x= element_text(size=fontsize.x, face = 'bold'),
        axis.text.y= element_text(size=fontsize.y, face = 'bold'), 
        axis.title =element_text(size=10),
        legend.title = element_blank()) +ggtitle(title)
}


# Application 
gsea.res4 = perform_GSEA(res = res4, ref = hallmark) 

filtered_gsea4 = gsea.res4 %>% mutate(sig= ifelse(pvalue <= 0.05,"p value <= 0.05", "p value > 0.05"))
```

#### GSEA NES Plot 1
```{r, fig.width=8, fig.height=7}
t= paste0("DEGs from ",tret, " / ", cont )

gsea_nes_plot_pval(filtered_gsea4, title = t, fontsize.x = 8, fontsize.y = 8)
```

#### GSEA NES Plot 2
```{r, fig.width=8, fig.height=7}
t= paste0("DEGs from ",tret, " / ", cont )

gsea_nes_plot(filtered_gsea4, title = t, fontsize.x = 8, fontsize.y = 8)
```

### DS-1062a vs IgG {.tabset}

```{r}
library(DESeq2)
count.mtx = counts
cont = "IgG"
tret = "DS1062a"
cols = c(c(7:9), c(10:12))
cond = c(rep(cont,3),
         rep(tret,3))
count.mtx = count.mtx[, cols]
# Generate info table
info <- data.frame(matrix(nrow = ncol(count.mtx), ncol = 2))
colnames(info) <- c('sample', 'cond')
info$sample <- colnames(count.mtx)
info$cond <- cond

# DESeq
dds <- DESeqDataSetFromMatrix(count.mtx, info, ~ cond)
dds <- DESeq(dds)
res <- results(dds)

res <- data.frame(res)
res5 <- res
```

#### Volcano plot (version1)
```{r}
res5 = add_DE_information(res = res5)

res5 %>% 
  ggplot(aes(log2FoldChange, -log10(padj))) + 
  geom_point(size=1, alpha=0.5) + 
  scale_color_manual(values = c("red3","royalblue3","grey"), guide = FALSE) +
  theme_classic() +
  geom_vline(xintercept = c(-log2(fc),log2(fc)), color='grey') +
  geom_hline(yintercept = -log10(pval),color='grey') 
```

#### Volcano plot (version2)
```{r, fig.width=8, fig.height=5}
plot_DE_information(res5, fc = 1.5, pval = 0.05, p_col = 'padj')
plot_DE_information(res5, fc = 1.5, pval = 0.05, p_col = 'pvalue')
```

#### Volcano plot (version3)
```{r, fig.width=8, fig.height=5}

p1 = plot_DE_information(res5, fc = 1.5, pval = 0.05, p_col = 'padj')

p2 = plot_DE_information(res5, fc = 1.5, pval = 0.05, p_col = 'pvalue')

p1 + scale_y_sqrt()
p2 + scale_y_sqrt()
```


#### DEG table
```{r}
res5 %>% DT::datatable(extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=10))
```


```{r}
# Application 
gsea.res5 = perform_GSEA(res = res5, ref = hallmark) 

filtered_gsea5 = gsea.res5 %>% mutate(sig= ifelse(pvalue <= 0.05,"p value <= 0.05", "p value > 0.05"))
```

#### GSEA NES Plot 1
```{r, fig.width=8, fig.height=7}
t= paste0("DEGs from ",tret, " / ", cont )

gsea_nes_plot_pval(filtered_gsea5, title = t, fontsize.x = 8, fontsize.y = 8)
```

#### GSEA NES Plot 2
```{r, fig.width=8, fig.height=7}
t= paste0("DEGs from ",tret, " / ", cont )

gsea_nes_plot(filtered_gsea5, title = t, fontsize.x = 8, fontsize.y = 8)
```

## DEGs comparison (by Adjusted p value){.tabset}

### UP-regulated genes

```{r}
# Read res list
degs = list(res4=res4,
            res5=res5)
names(degs) = c("DS-1062a/Dxd",
                "DS-1062a/IgG-Dxd")
```

```{r}
column = "DE_padj"
# 'UP'으로 표시된 유전자 이름을 추출하여 새로운 리스트 생성
selected_genes <- lapply(degs, function(df) {
  df %>% filter(!!sym(column) == "UP") %>% rownames()
})

# degs 리스트의 각 요소와 동일한 이름을 새 리스트에 할당
names(selected_genes) <- names(degs)
```

```{r}
ven_list = selected_genes

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```

```{r}
plot(ven_out)
```

```{r}
plot(ven_out, type = "upset")
```

```{r}
ven_out@result %>% DT::datatable()
```


### DN-regulated genes
```{r}
column = "DE_padj"
# 'DN'으로 표시된 유전자 이름을 추출하여 새로운 리스트 생성
selected_genes <- lapply(degs, function(df) {
  df %>% filter(!!sym(column) == "DN") %>% rownames()
})

# degs 리스트의 각 요소와 동일한 이름을 새 리스트에 할당
names(selected_genes) <- names(degs)
```

```{r}
ven_list = selected_genes

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```

```{r}
plot(ven_out)
```

```{r}
plot(ven_out, type = "upset")
```

```{r}
ven_out@result %>% DT::datatable()
```


```{r}
## save res files
res4 %>% write.csv(paste0(dir,"data/30342/res4.24.06.25.csv"))
res5 %>% write.csv(paste0(dir,"data/30342/res5.24.06.25.csv"))
```


## DEGs comparison (by p value){.tabset}

### UP-regulated genes 

```{r}
column <- "DE_pvalue"

# Extract gene names labeled as 'UP' into a new list
up_genes <- lapply(degs, function(df) {
  df$gene <- rownames(df)
  df %>% filter(!!sym(column) == "UP") %>% select(gene) %>% pull()
})
```


```{r}
ven_list = up_genes

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```

```{r}
plot(ven_out) 
```

```{r}
plot(ven_out, type = "upset") 
```

```{r}
ven_out@result %>% DT::datatable()
```


### DN-regulated genes 
```{r}
column = "DE_pvalue"
# 'DN'으로 표시된 유전자 이름을 추출하여 새로운 리스트 생성
selected_genes <- lapply(degs, function(df) {
  df %>% filter(!!sym(column) == "DN") %>% rownames()
})

# degs 리스트의 각 요소와 동일한 이름을 새 리스트에 할당
names(selected_genes) <- names(degs)
```


```{r}
ven_list = selected_genes

## draw venn diagram
ven_out <- VennDetail::venndetail(ven_list)
```

```{r}
plot(ven_out) 
```

```{r}
plot(ven_out, type = "upset") 
```

```{r}
ven_out@result %>% DT::datatable()
```