---
title: "2024 Daiichi DS1062a"
subtitle: "bulk RNA-seq"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(DT)
```

### Introduction

Paweletz lab  
2024.06.10
<br> 

<!-- [**Link to Daiichi DS1062a scRNA-seq**](../../2024_Daiichi_DXD/rmd/Daiichi_2024_index.html) -->


### Experimental Conditions  

- **Patient Groups:** 30342 and 30348  
- **Treatment Conditions:**  
  - **M (Control):** DMSO  
  - **D:** Dxd (deruxtecan, topoisomerase inhibitor) 1 µM  
  - **I:** IgG Dxd (non-specific gene-targeting Dxd) 30 µg/ml  
  - **A:** DS-1062a (datopotamab deruxtecan, TROP2-targeting Dxd) 30 µg/ml
- **Sequencing type:** bulk RNA-seq  


### Data Information  

#### Table  
| Sample ID   | Treatment |
|-------------|-----------|
| 30342-1.1   | DMSO      |
| 30342-1.2   | DMSO      |
| 30342-1.3   | DMSO      |
| 30342-2.1   | Dxd       |
| 30342-2.2   | Dxd       |
| 30342-2.3   | Dxd       |
| 30342-3.1   | IgG-DxD   |
| 30342-3.2   | IgG-DxD   |
| 30342-3.3   | IgG-DxD   |
| 30342-4.1   | DS1062a   |
| 30342-4.2   | DS1062a   |
| 30342-4.3   | DS1062a   |
| 30348-1.1   | DMSO      |
| 30348-1.2   | DMSO      |
| 30348-1.3   | DMSO      |
| 30348-2.1   | Dxd       |
| 30348-2.2   | Dxd       |
| 30348-2.3   | Dxd       |
| 30348-3.1   | IgG-DxD   |
| 30348-3.2   | IgG-DxD   |
| 30348-3.3   | IgG-DxD   |
| 30348-4.1   | DS1062a   |
| 30348-4.2   | DS1062a   |
| 30348-4.3   | DS1062a   |

**30342 only**  

| Sample ID | DMSO | Dxd | IgG-DxD | DS1062a |
|-----------|------|-----|---------|---------|
| 30342-1.1 | DMSO |     |         |         |
| 30342-1.2 | DMSO |     |         |         |
| 30342-1.3 | DMSO |     |         |         |
| 30342-2.1 |      | Dxd |         |         |
| 30342-2.2 |      | Dxd |         |         |
| 30342-2.3 |      | Dxd |         |         |
| 30342-3.1 |      |     | IgG-DxD |         |
| 30342-3.2 |      |     | IgG-DxD |         |
| 30342-3.3 |      |     | IgG-DxD |         |
| 30342-4.1 |      |     |         | DS1062a |
| 30342-4.2 |      |     |         | DS1062a |
| 30342-4.3 |      |     |         | DS1062a |

**30348 only**  

| Sample ID | DMSO | Dxd | IgG-DxD | DS1062a |
|-----------|------|-----|---------|---------|
| 30348-1.1 | DMSO |     |         |         |
| 30348-1.2 | DMSO |     |         |         |
| 30348-1.3 | DMSO |     |         |         |
| 30348-2.1 |      | Dxd |         |         |
| 30348-2.2 |      | Dxd |         |         |
| 30348-2.3 |      | Dxd |         |         |
| 30348-3.1 |      |     | IgG-DxD |         |
| 30348-3.2 |      |     | IgG-DxD |         |
| 30348-3.3 |      |     | IgG-DxD |         |
| 30348-4.1 |      |     |         | DS1062a |
| 30348-4.2 |      |     |         | DS1062a |
| 30348-4.3 |      |     |         | DS1062a |

#### Raw fastq files backup 

**rc-stor15 location**   
/mnt/centers/bxdx/srp68/raw_data/Daiichi_DS1062a_bulk/fastq  

<br><br>

#### Dropbox Link  

[**Dropbox folder : 2024_Daiichi_DS1062a_bulk**](https://www.dropbox.com/scl/fo/led84hq7i60c5ofvav69h/AHsmEzDqtoOKrdcieT7MJ44?rlkey=0f0cp5siv8m60iy3ikyz7ameq&dl=0)

#### MultiQC file 

[**Report Link**](multiqc_report.html)<br>
<br><br>


