--- 
title: "MT30271 MT30273 comparison analysis"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::readthedown: 
    code_folding: hide 
    number_sections: TRUE
--- 


<style> 
#TOC { 
  top: 1%; 
  opacity: 0.5; 
} 
#TOC:hover { 
  opacity: 1; 
} 
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
```

**R Markdown**

This is a document for subset data of MT30271 and MT30273.

The details regarding MT30217 and MT30273 are currently confidential.

<br>
<hr>
<br>

# Import preprocessed data
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
obj.srt = readRDS(paste0(dir,'rds/MT30271.30273.23.07.31.rds'))
```

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 20000. 

mitochondrial cutoff : 25. 

other cutoffs : NA. 

doublet removal (scrublet) : Y. 

regression: ribosomal genes. 

# Gene expression on UMAP

## Cell type marker genes
```{r}
data <- data.frame(
  Category = c("B", "Macrophage", "DC", "Fibroblast", "Tcell", "NK", "RBC", "Epithelial", "Tumor", "Monocyte"),
  Genes = c("CD79A, IGHM, CD79B, MS4A1",
            "CD14, VSIG4, C1QB, C1QA, APOE, SPP1, CD83",
            "ITGAX, ITGAM, CD1C,  FLT3, HLA-DRA, HLA-DRB1",
            "COL3A1, COL4A1, COL4A2, S100A16, UGDH",
            "TCF7, CD3G, CD3E, CD3D, CD4, CD28, BCL11B, CD8A, CD8B, ICOS",
            "GNLY, KLRD1, KLRB1, KLRC1",
            "HBB",
            "KRT18, KRT8, KRT18, KRT19",
            "CTLA4, HAVCR2, EGFR, KRAS, MYC, CD274",
            "CD14, CD68, CCR2, CSF1R")
)

DT::datatable(data, caption = 'Cell type marker genes') 
```

```{r, echo=F}
genes_vector <- c("CD79A", "IGHM", "CD79B", "MS4A1",
                  "CD14", "VSIG4", "C1QB", "C1QA", "APOE", "SPP1", "CD83",
                  "ITGAX", "ITGAM", "CD1C", "FLT3", "HLA-DRA", "HLA-DRB1",
                  "COL3A1", "COL4A1", "COL4A2", "S100A16", "UGDH",
                  "TCF7", "CD3G", "CD3E", "CD3D", "CD4", "CD28", "BCL11B", "CD8A", "CD8B", "ICOS",
                  "GNLY", "KLRD1", "KLRB1", "KLRC1",
                  "HBB",
                  "KRT18", "KRT8", "KRT18", "KRT19",
                  "CTLA4", "HAVCR2", "EGFR", "KRAS", "MYC", "CD274",
                  "CD14", "CD68", "CCR2", "CSF1R")
```


## B cell type
```{r, fig.height=6, fig.width=8}
genes = c("CD79A", "IGHM", "CD79B", "MS4A1")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 2)
```

## Macrophage type
```{r, fig.height=9, fig.width=10}
genes = c("CD14", "VSIG4", "C1QB", "C1QA", "APOE", "SPP1", "CD83")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 3)
```

## DC type
```{r, fig.height=6, fig.width=10}
genes = c("ITGAX", "ITGAM", "CD1C", "FLT3", "HLA-DRA", "HLA-DRB1")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 3)
```

## Fibroblast type
```{r, fig.height=6, fig.width=10}
genes = c("COL3A1", "COL4A1", "COL4A2", "S100A16", "UGDH")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 3)
```

## Tcell type
```{r, fig.height=9, fig.width=15}
genes = c("TCF7", "CD3G", "CD3E", "CD3D", "CD4", "CD28", "BCL11B", "CD8A", "CD8B", "ICOS")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 4)
```

## NK type
```{r, fig.height=6, fig.width=8}
genes = c("GNLY", "KLRD1", "KLRB1", "KLRC1")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 2)
```

## RBC type
```{r, fig.height=3, fig.width=4}
genes = c("HBB")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 1)
```

## Epithelial cell type
```{r, fig.height=6, fig.width=8}
genes = c("KRT18", "KRT8", "KRT18", "KRT19")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 2)
```

## Tumor type
```{r, fig.height=6, fig.width=10}
genes = c("CTLA4", "HAVCR2", "EGFR", "KRAS", "MYC", "CD274")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 3)
```

## Monocyte type
```{r, fig.height=6, fig.width=8}
genes = c("CD14", "CD68", "CCR2", "CSF1R")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 2)
```
