---
title: "MT30217 MT30273"
author: "Sung Rye Park"
date: "2023-08-10"
output: 
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, fig.align = "center", message=F, warning=F)
```

## R Markdown

This is a document for subset data of MT30217 and MT30273.

The details regarding MT30217 and MT30273 are currently confidential.



```{r, echo=F}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```

#### Import the processed data
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
obj.srt = readRDS(paste0(dir,'rds/MT30271.30273.23.07.31.rds'))
```

#### pre-processing information
UMI cutoff : 500~ 20000. 

mitochondrial cutoff : 25. 

other cutoffs : NA. 

doublet removal (scrublet) : Y. 

regression: ribosomal genes. 
   
   
---   
#### Visualization
UMAP  
```{r, fig.align='center', echo=TRUE}
DimPlot(obj.srt, group.by = 'orig.ident')
DimPlot(obj.srt, group.by = 'res')
DimPlot(obj.srt, group.by = 'sample')
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 2)
DimPlot(obj.srt, group.by = 'RNA_snn_res.0.2')
DimPlot(obj.srt, group.by = 'RNA_snn_res.0.2', label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


#### Display the number of cells in each sample 
```{r, echo=F}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable()
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable()
obj.srt@meta.data %>% select(RNA_snn_res.0.2) %>% table() %>% data.frame() %>% DT::datatable()
obj.srt@meta.data %>% select(RNA_snn_res.0.2, res) %>% table() %>% data.frame() %>% tidyr::spread(res, Freq) %>% DT::datatable()
obj.srt@meta.data %>% select(RNA_snn_res.0.2, res, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable()
```

```{r, echo=FALSE}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
all.markers = read.csv(paste0(dir,'data/MT30271.MT30273/mks/res.0.2.all.markers.23.07.31.csv'))
## top marker genes table 
mks= all.markers
df=mks$cluster %>% table() %>% data.frame()
df[,'top_genes'] =''
for(i in 1:length(unique(mks$cluster))){
  if(df[i,]$Freq >= 30){
    cl = unique(mks$cluster)[i]
    df[i,'top_genes'] = mks %>% filter(cluster== cl) %>% arrange(desc(avg_log2FC)) %>% head(30) %>% select(gene) %>% pull() %>% paste0(collapse = ',')
  }else{
    cl = unique(mks$cluster)[i]
    df[i,'top_genes'] = mks %>% filter(cluster== cl) %>% arrange(desc(avg_log2FC)) %>% select(gene) %>% pull() %>% paste0(collapse = ',')
  }
}
```
  
Top 30 genes table
```{r}
DT::datatable(df) 
```

cell type guess

```{r, echo=F}
data <- data.frame(
  Category = c("B", "Macrophage", "DC", "Fibroblast", "Tcell", "NK", "RBC", "Epithelial", "Tumor", "Monocyte"),
  Genes = c("CD79A, IGHM, CD79B, MS4A1",
            "CD14, VSIG4, C1QB, C1QA, APOE, SPP1, CD83",
            "ITGAX, ITGAM, CD1C,  FLT3, HLA-DRA, HLA-DRB1",
            "COL3A1, COL4A1, COL4A2, S100A16, UGDH",
            "TCF7, CD3G, CD3E, CD3D, CD4, CD28, BCL11B, CD8A, CD8B, ICOS",
            "GNLY, KLRD1, KLRB1, KLRC1",
            "HBB",
            "KRT18, KRT8, KRT18, KRT19",
            "CTLA4, HAVCR2, EGFR, KRAS, MYC, CD274",
            "CD14, CD68, CCR2, CSF1R")
)

print(data)

```


```{r, echo=F}
genes_vector <- c("CD79A", "IGHM", "CD79B", "MS4A1",
                  "CD14", "VSIG4", "C1QB", "C1QA", "APOE", "SPP1", "CD83",
                  "ITGAX", "ITGAM", "CD1C", "FLT3", "HLA-DRA", "HLA-DRB1",
                  "COL3A1", "COL4A1", "COL4A2", "S100A16", "UGDH",
                  "TCF7", "CD3G", "CD3E", "CD3D", "CD4", "CD28", "BCL11B", "CD8A", "CD8B", "ICOS",
                  "GNLY", "KLRD1", "KLRB1", "KLRC1",
                  "HBB",
                  "KRT18", "KRT8", "KRT18", "KRT19",
                  "CTLA4", "HAVCR2", "EGFR", "KRAS", "MYC", "CD274",
                  "CD14", "CD68", "CCR2", "CSF1R")

# Reverse the order of 'ident' levels
obj.srt@meta.data$rev.cluster <- factor(obj.srt@meta.data$RNA_snn_res.0.2, 
                                        levels = rev(levels(obj.srt@meta.data$RNA_snn_res.0.2)))
```

  
```{r, echo=F}
VlnPlot(obj.srt, features = genes_vector, group.by = "rev.cluster",
        stack = T, flip = F) + NoLegend()
```

```{r}
data <- data.frame(
  cluster = c(
    "clustere0", "clustere1", "clustere2", "clustere3", "clustere4",
    "clustere5", "clustere6", "clustere7", "clustere8", "clustere9",
    "clustere10", "clustere11", "clustere12", "clustere13", "clustere14",
    "clustere15", "clustere16"
  ),
  type = c(
    "T cell", "T cell", "Macrophage", "T cell2", "Fibroblast tumor-associated",
    "Epithelial", "NK", "Macrophage", "Fibroblast", "Fibroblast",
    "UKN", "T cell-tumor associated", "mono/macrophage", "B cell",
    "UKN", "mono/macrophage-tumor associated", "mono/macrophage-tumor associated2"
  )
)
print(data)
```

  
```{r}
# add cell type based on cluster resolution 0.2
type_vector <- c(
  "T cell1", "T cell2", "Macrophage1", "T cell2", "Fibroblast tumor-associated",
  "Epithelial", "NK", "Macrophage2", "Fibroblast1", "Fibroblast2",
  "UKN1", "T cell-tumor associated", "mono/macrophage", "B cell",
  "UKN2", "mono/macrophage-tumor associated 1", "mono/macrophage-tumor associated 2"
)

obj.srt@meta.data$cell_type = factor(obj.srt@meta.data$RNA_snn_res.0.2)
levels(obj.srt@meta.data$cell_type) = type_vector
```

```{r}
DimPlot(obj.srt, group.by = 'cell_type')
DimPlot(obj.srt, group.by = 'cell_type', label = T, label.box = T, label.size = 3) + NoLegend()
```

```{r}
# add main cell type based on cluster resolution 0.2
types <- c(
  "T cell", "T cell", "macrophage", "T cell", "Fibroblast-tumor-associated",
  "Epithelial cell", "NK", "macrophage", "fibroblast", "fibroblast",
  "UKN", "T cell-tumor associated", "mono/macrophage", "B cell",
  "UKN", "mono/macrophage-tumor associated", "mono/macrophage-tumor associated"
)
obj.srt@meta.data$main_cell_type = factor(obj.srt@meta.data$RNA_snn_res.0.2)
levels(obj.srt@meta.data$main_cell_type) = types

```

```{r}
DimPlot(obj.srt, group.by = 'main_cell_type')
DimPlot(obj.srt, group.by = 'main_cell_type', label = T, label.box = T, label.size = 3) + NoLegend()
```

```{r, fig.height=15, fig.width=7}
obj.srt@reductions$umap@cell.embeddings %>% data.frame() %>% 
  ggplot(aes(UMAP_1, UMAP_2, color=obj.srt$main_cell_type)) + geom_point(size=0.2) +
  facet_wrap(.~obj.srt$sample, ncol = 1) +
  theme_classic() +guides(colour = guide_legend(override.aes = list(size=5))) +
  labs(color = "Main Cell Type")

```

```{r, echo=FALSE}
## sankey plot 
## sankey plot 
library(ggsankey)
library(ggplot2)
library(dplyr)

data =obj.srt@meta.data[, c('orig.ident','RNA_snn_res.0.2','main_cell_type')]

# Function to create a Sankey plot
create_sankey_plot <- function(data, column_names, title, show_labels = FALSE){
  df <- data %>%
    make_long(!!column_names[1], !!column_names[2], !!column_names[3])
  
  dagg <- df %>%
    group_by(node) %>%
    tally()
  
  df2 <- merge(df, dagg, by.x = 'node', by.y = 'node', all.x = TRUE)
  
  pl <- ggplot(df2, aes(x = x,
                        next_x = next_x,
                        node = node,
                        next_node = next_node,
                        fill = factor(node),
                        label = paste0(node," n=", n))
  )
  
  pl <- pl + geom_sankey(flow.alpha = 0.5, color = "gray40", show.legend = show_labels)
  if (show_labels) {
    pl <- pl + geom_sankey_label(size = 3, color = "white", fill = "gray40", hjust = 1)  # Change hjust value to 1 (right-aligned)
  }
  
  pl <- pl + theme_bw()
  pl <- pl + theme(legend.position = "none")
  pl <- pl + theme(axis.title = element_blank(),
                   axis.text.y = element_blank(),
                   axis.ticks = element_blank(),
                   panel.grid = element_blank())
  pl <- pl + scale_fill_viridis_d(option = "inferno")
  pl <- pl + labs(title = "Sankey diagram")
  pl <- pl + labs(subtitle = "cluster distribution")
  
  pl <- pl + labs(fill = 'Nodes')
  
  return(pl)
}

# Assuming obj.srt@meta.data is your data and you want to use specific columns for the Sankey plot
column_names <- c('orig.ident', 'RNA_snn_res.0.2','main_cell_type')

# Create the Sankey plot using the specified columns
sankey_plot <- create_sankey_plot(data = obj.srt@meta.data, column_names, 
                                  title = "MT30271_MT30273", show_labels = TRUE)
```

Sankey plot
```{r,fig.height=6, fig.width=10}
# Display the plot
print(sankey_plot)
```
