---
title: "MT30271 MT30273 comparison analysis"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::readthedown: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 4
    toc_float: true
--- 


<style> 
#TOC { 
  top: 1%; 
  opacity: 0.5; 
} 
#TOC:hover { 
  opacity: 1; 
} 
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
```


**R Markdown**

This is a document for subset data of MT30217 and MT30273.

The details regarding MT30217 and MT30273 are currently confidential.

<br>
<hr>
<br>

# Import preprocessed data
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
obj.srt = readRDS(paste0(dir,'rds/MT30271.30273.23.07.31.rds'))
```

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 20000. 

mitochondrial cutoff : 25. 

other cutoffs : NA. 

doublet removal (scrublet) : Y. 

regression: ribosomal genes. 

---

# Number of cells in main cell types

```{r}
obj.srt@meta.data %>% select(main_cell_type, res, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by main cell type and R/NR and treatment')
```


** Define functions for downstream analysis**
```{r, echo=F}
#######################
# Define functions
set_compare_column <- function(obj.srt, filter1, filter1_selection, filter2, filter2_selection, filter3, selected_groups) {
  rows1 <- obj.srt@meta.data %>% 
    filter(get(filter1) == filter1_selection) %>% 
    filter(get(filter2) == filter2_selection) %>% 
    filter(get(filter3) %in% selected_groups[1]) %>% rownames()
  
  rows2 <- obj.srt@meta.data %>% 
    filter(get(filter1) == filter1_selection) %>% 
    filter(get(filter2) == filter2_selection) %>% 
    filter(get(filter3) %in% selected_groups[2]) %>% rownames()
  
  obj.srt@meta.data <- obj.srt@meta.data %>%
    mutate(compare = case_when(
      rownames(.) %in% rows1 ~ 'g1',
      rownames(.) %in% rows2 ~ 'g2',
      TRUE ~ 'other'
    ))
  
  return(obj.srt)
}
DEG.function = function(obj.srt,
                        filter1 = filter1_selection,
                        groups = selected_groups){
  # add compare to the ident
  Idents(obj.srt) = 'compare'
  logfc=log2(1)
  mks =FindMarkers(obj.srt, ident.1 = 'g2', ident.2 = 'g1', 
                   logfc.threshold = logfc)
  pval=0.05
  fc=1.2
  mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                                 ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))
  mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))
  mks$gene = rownames(mks)
  mks =mks %>% mutate(labels= ifelse(DE == 'UP', gene, ifelse(DE=='DN',gene,'other')))
  mks =mks %>% arrange(desc(avg_log2FC))
  mks[,'main_cell_type']= filter1
  mks[,'comparison'] = paste0(groups, collapse = '_')
  return(mks)
}
plot_volcano_text <- function(mks, title) {
  if (length(unique(mks$DE)) == 3) {
    # Set DE levels and add gene names
    mks$DE <- factor(mks$DE, levels = c('UP', 'DN', 'no_sig'))
    mks$gene <- rownames(mks)
    
    # Set fold change threshold
    fc <- 1.2
    
    # Create the ggplot plot
    p <- mks %>%
      mutate(labels = ifelse(labels == 'other', '', labels)) %>%
      ggplot(aes(avg_log2FC, -log10(p_val_adj), color = DE)) + 
      geom_point(size = 1, alpha = 0.5) + 
      scale_color_manual(values = c('red', 'blue', 'grey')) +
      theme_classic() +
      geom_vline(xintercept = c(-log2(fc), log2(fc)), color = 'grey') +
      geom_hline(yintercept = -log10(0.05), color = 'grey') +
      guides(colour = guide_legend(override.aes = list(size = 5))) +
      ggtitle(title) +
      # Use geom_text with hjust argument to adjust label position
      geom_text(aes(label = labels), size = 2.5, show.legend = FALSE, hjust = 0, nudge_x = 0.01) +
      ggeasy::easy_center_title() ## to center title
    
    # Print the plot
    print(p)
  } else if (length(unique(mks$DE)) == 2 && all(unique(mks$DE) %in% c('UP', 'no_sig'))) {
    # Set DE levels and add gene names
    mks$DE <- factor(mks$DE, levels = c('UP', 'no_sig'))
    mks$gene <- rownames(mks)
    
    # Set fold change threshold
    fc <- 1.2
    
    # Create the ggplot plot
    p <- mks %>%
      mutate(labels = ifelse(labels == 'other', '', labels)) %>%
      ggplot(aes(avg_log2FC, -log10(p_val_adj), color = DE)) + 
      geom_point(size = 1, alpha = 0.5) + 
      scale_color_manual(values = c('red', 'grey')) +
      theme_classic() +
      geom_vline(xintercept = c(-log2(fc), log2(fc)), color = 'grey') +
      geom_hline(yintercept = -log10(0.05), color = 'grey') +
      guides(colour = guide_legend(override.aes = list(size = 5))) +
      ggtitle(title) +
      # Use geom_text with hjust argument to adjust label position
      geom_text(aes(label = labels), size = 2.5, show.legend = FALSE, hjust = 0, nudge_x = 0.01) +
      ggeasy::easy_center_title() ## to center title
    
    # Print the plot
    print(p)
  } else if (length(unique(mks$DE)) == 2 && all(unique(mks$DE) %in% c('DN', 'no_sig'))) {
    # Set DE levels and add gene names
    mks$DE <- factor(mks$DE, levels = c('DN', 'no_sig'))
    mks$gene <- rownames(mks)
    
    # Set fold change threshold
    fc <- 1.2
    
    # Create the ggplot plot
    p <- mks %>%
      mutate(labels = ifelse(labels == 'other', '', labels)) %>%
      ggplot(aes(avg_log2FC, -log10(p_val_adj), color = DE)) + 
      geom_point(size = 1, alpha = 0.5) + 
      scale_color_manual(values = c('blue', 'grey')) +
      theme_classic() +
      geom_vline(xintercept = c(-log2(fc), log2(fc)), color = 'grey') +
      geom_hline(yintercept = -log10(0.05), color = 'grey') +
      guides(colour = guide_legend(override.aes = list(size = 5))) +
      ggtitle(title) +
      # Use geom_text with hjust argument to adjust label position
      geom_text(aes(label = labels), size = 2.5, show.legend = FALSE, hjust = 0, nudge_x = 0.01) +
      ggeasy::easy_center_title() ## to center title
    
    # Print the plot
    print(p)
  }
}
```

# Find DEGs and plot them in volcanoplot


## fibroblast 

### fibroblast & Responder & CTL vs TAK

** Filtering parameters**
```{r}
# filtering parameters
filter_data <- data.frame(
  filter1 = 'main_cell_type',
  filter1_selection = 'fibroblast',
  filter2 = 'res',
  filter2_selection = 'R',
  filter3 = 'sample',
  selected_groups = c('CTL', 'TAK')
)

filter_data %>% gt::gt()

```


```{r}
# filtering parameters
filter1= 'main_cell_type'
filter1_selection= 'fibroblast'
filter2= 'res'
filter2_selection= 'R'
filter3= 'sample'
selected_groups= c('CTL','TAK')

# Define groups to compare
obj.srt <- set_compare_column(obj.srt,
                              filter1 = filter1,
                              filter1_selection = filter1_selection,
                              filter2 = filter2,
                              filter2_selection = filter2_selection,
                              filter3 = filter3,
                              selected_groups = selected_groups)
# Check the table of the 'compare' column
# table(obj.srt@meta.data$compare) 

# Find DEGs
DEG.out= DEG.function(obj.srt,
                      filter1 = filter1_selection,
                      groups = selected_groups)

```


#### Volcanoplot

```{r}
# Plot volcano plot for DEGs
plot_volcano_text(mks = DEG.out, title=paste0(filter1_selection, ' ',
                                              filter2_selection, ' ',
                                              paste0(selected_groups, collapse = ' vs ')))
```



#### Up-regulated genes

```{r}
data.frame("UP-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'UP') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```



#### DOWN-regulated genes

```{r}
data.frame("DN-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'DN') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```






### fibroblast & Responder & CTL vs COM

** Filtering parameters**
```{r}
# filtering parameters
filter_data <- data.frame(
  filter1 = 'main_cell_type',
  filter1_selection = 'fibroblast',
  filter2 = 'res',
  filter2_selection = 'R',
  filter3 = 'sample',
  selected_groups = c('CTL', 'COM')
)

filter_data %>% gt::gt()

```


```{r}
# filtering parameters
filter1= 'main_cell_type'
filter1_selection= 'fibroblast'
filter2= 'res'
filter2_selection= 'R'
filter3= 'sample'
selected_groups= c('CTL','COM')

# Define groups to compare
obj.srt <- set_compare_column(obj.srt,
                              filter1 = filter1,
                              filter1_selection = filter1_selection,
                              filter2 = filter2,
                              filter2_selection = filter2_selection,
                              filter3 = filter3,
                              selected_groups = selected_groups)
# Check the table of the 'compare' column
# table(obj.srt@meta.data$compare) 

# Find DEGs
DEG.out= DEG.function(obj.srt,
                      filter1 = filter1_selection,
                      groups = selected_groups)

```


#### Volcanoplot

```{r}
# Plot volcano plot for DEGs
plot_volcano_text(mks = DEG.out, title=paste0(filter1_selection, ' ',
                                              filter2_selection, ' ',
                                              paste0(selected_groups, collapse = ' vs ')))
```



#### Up-regulated genes

```{r}
data.frame("UP-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'UP') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```



#### DOWN-regulated genes

```{r}
data.frame("DN-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'DN') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```





### fibroblast & Non Responder & CTL vs TAK

** Filtering parameters**
```{r}
# filtering parameters
filter_data <- data.frame(
  filter1 = 'main_cell_type',
  filter1_selection = 'fibroblast',
  filter2 = 'res',
  filter2_selection = 'NR',
  filter3 = 'sample',
  selected_groups = c('CTL', 'TAK')
)

filter_data %>% gt::gt()

```


```{r}
# filtering parameters
filter1= 'main_cell_type'
filter1_selection= 'fibroblast'
filter2= 'res'
filter2_selection= 'NR'
filter3= 'sample'
selected_groups= c('CTL','TAK')

# Define groups to compare
obj.srt <- set_compare_column(obj.srt,
                              filter1 = filter1,
                              filter1_selection = filter1_selection,
                              filter2 = filter2,
                              filter2_selection = filter2_selection,
                              filter3 = filter3,
                              selected_groups = selected_groups)
# Check the table of the 'compare' column
# table(obj.srt@meta.data$compare) 

# Find DEGs
DEG.out= DEG.function(obj.srt,
                      filter1 = filter1_selection,
                      groups = selected_groups)

```


#### Volcanoplot

```{r}
# Plot volcano plot for DEGs
plot_volcano_text(mks = DEG.out, title=paste0(filter1_selection, ' ',
                                              filter2_selection, ' ',
                                              paste0(selected_groups, collapse = ' vs ')))
```



#### Up-regulated genes

```{r}
data.frame("UP-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'UP') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```



#### DOWN-regulated genes

```{r}
data.frame("DN-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'DN') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```






### fibroblast & Non Responder & CTL vs COM

** Filtering parameters**
```{r}
# filtering parameters
filter_data <- data.frame(
  filter1 = 'main_cell_type',
  filter1_selection = 'fibroblast',
  filter2 = 'res',
  filter2_selection = 'NR',
  filter3 = 'sample',
  selected_groups = c('CTL', 'COM')
)

filter_data %>% gt::gt()

```


```{r}
# filtering parameters
filter1= 'main_cell_type'
filter1_selection= 'fibroblast'
filter2= 'res'
filter2_selection= 'NR'
filter3= 'sample'
selected_groups= c('CTL','COM')

# Define groups to compare
obj.srt <- set_compare_column(obj.srt,
                              filter1 = filter1,
                              filter1_selection = filter1_selection,
                              filter2 = filter2,
                              filter2_selection = filter2_selection,
                              filter3 = filter3,
                              selected_groups = selected_groups)
# Check the table of the 'compare' column
# table(obj.srt@meta.data$compare) 

# Find DEGs
DEG.out= DEG.function(obj.srt,
                      filter1 = filter1_selection,
                      groups = selected_groups)

```


#### Volcanoplot

```{r}
# Plot volcano plot for DEGs
plot_volcano_text(mks = DEG.out, title=paste0(filter1_selection, ' ',
                                              filter2_selection, ' ',
                                              paste0(selected_groups, collapse = ' vs ')))
```



#### Up-regulated genes

```{r}
data.frame("UP-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'UP') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```



#### DOWN-regulated genes

```{r}
data.frame("DN-regulated genes" = (DEG.out %>% 
                                     filter(DE == 'DN') %>% 
                                     select(gene) %>% 
                                     pull() %>% 
                                     paste0(collapse = ',')),
           row.names = 'fibroblast') %>% DT::datatable(options = list(scrollX=T))
```



