---
title: "2024 Daiichi DXD"
subtitle: "Initial Report: 30342+30348"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
## load required packages
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
```

<br>

```{r, eval=FALSE}
dir= "~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD/"
obj.342 = readRDS(paste0(dir,"rds/P30342.init.24.05.06.rds"))
obj.348 = readRDS(paste0(dir,"rds/P30348.init.24.05.07.rds"))

# merge two dataset 
# url (https://satijalab.org/seurat/archive/v4.3/merge)
mg = merge(obj.342, y = obj.348, add.cell.ids = c("P30342", "P30348"), project = "Daiichi")

mg %>% saveRDS(paste0(dir,"rds/P30342_30348_merged.24.05.07.rds"))
```


```{r}
dir= "~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD/"
obj.srt = readRDS(paste0(dir,"rds/P30342_30348_merged.24.05.07.rds"))
```


```{r, eval=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 2000, n_pcs = 20, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.2), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, 
                                  selection.method = 'vst', 
                                  nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- NormalizeData(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, 
                    features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj_merged)
```



**Number of total cells** 
```{r}
obj.srt@meta.data %>% nrow()
```

**Number of cells in each condition**
```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("sample","Number of cells"))
```

```{r, eval=FALSE}
obj.srt@meta.data %>%
  group_by(orig.ident) %>%
  summarise(across(where(is.numeric), list(mean = mean, median = median, sd = sd), na.rm = TRUE)) %>% DT::datatable()
```


**Useful information for filtering**  
[1. Introduction to Single-cell RNA-seq : Check Filtering section](https://hbctraining.github.io/scRNA-seq/lessons/04_SC_quality_control.html)<br>
[2. Common Considerations for Quality Control Filters for Single Cell RNA-seq Data: 10X Genomicx](https://www.10xgenomics.com/analysis-guides/common-considerations-for-quality-control-filters-for-single-cell-rna-seq-data)



# UMI  
```{r}
obj.srt@meta.data %>%
  ggplot(aes(x = orig.ident, y = nCount_RNA)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2, alpha = 0.1) +
  scale_y_log10()
```

# Mitochondrial content  
```{r}
obj.srt@meta.data %>%
  ggplot(aes(x = orig.ident, y = percent.mt)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2, alpha = 0.1) +
  scale_y_sqrt()
```

# UMI vs mitochondrial contents {.tabset}  

## version 1  
```{r, fig.width=16, fig.height=7}
obj.srt@meta.data %>% ggplot(aes(nCount_RNA, percent.mt)) +geom_point(size=0.5, alpha=0.5) +
  facet_wrap(.~orig.ident, ncol = 4) +scale_x_sqrt()
```


## version 2  
```{r, fig.width=16, fig.height=7}
obj.srt@meta.data %>% ggplot(aes(nCount_RNA, percent.mt)) +geom_point(size=0.3, alpha=0.5) +
  facet_wrap(.~orig.ident, ncol = 4) +scale_x_sqrt() +
  geom_vline(xintercept = c(100, 500, 1000), color="salmon") +
  geom_hline(yintercept = c(10, 25), color="salmon")
```

## version 3  
```{r, fig.width=16, fig.height=7}
obj.srt@meta.data %>%
  ggplot(aes(x = nCount_RNA, y = percent.mt)) +
  geom_point(size = 0.3, alpha = 0.5, color = "grey55") +
  stat_density_2d(aes(fill = ..level..), geom = "density_2d", color = "red", alpha=0.6) +
  facet_wrap(. ~ orig.ident, ncol = 4) +
  scale_fill_gradient(low = "grey", high = "red") +
  scale_x_sqrt() +
  geom_vline(xintercept = 100, color = "pink") +
  geom_vline(xintercept = 500, color = "pink") +
  geom_vline(xintercept = 1000, color = "pink") +
  geom_hline(yintercept = c(10, 25), color = "pink") +
  annotate("text", x = 100, y = max(obj.srt@meta.data$percent.mt) * 0.95, label = "100", color = "red", vjust = -1, size=3) +
  annotate("text", x = 500, y = max(obj.srt@meta.data$percent.mt) * 0.92, label = "500", color = "red", vjust = -1, size=3) +
  annotate("text", x = 1000, y = max(obj.srt@meta.data$percent.mt) * 0.89, label = "1000", color = "red", vjust = -1, size=3) +
  annotate("text", x = max(obj.srt@meta.data$nCount_RNA) * 0.95, y = 10, label = "10", color = "red", hjust = -0.1) +
  annotate("text", x = max(obj.srt@meta.data$nCount_RNA) * 0.95, y = 25, label = "25", color = "red", hjust = -0.1) +
  theme_minimal() 

```


# Number of cells with UMI & mitochondrial content cutoff {.tabset} 

## nCount_RNA >= 500 & percent.mt <= 25  
```{r}
obj.srt@meta.data %>% filter(predicted_doublet == "False") %>% filter(nCount_RNA >= 500 & percent.mt <= 25) %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("sample", "Number of cells"), caption = "nCount_RNA >= 500 & percent.mt <= 25")
```

## nCount_RNA >= 1000 & percent.mt <= 25  
```{r}
obj.srt@meta.data %>% filter(predicted_doublet == "False") %>% filter(nCount_RNA >= 1000 & percent.mt <= 25) %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("sample", "Number of cells"), caption = "nCount_RNA >= 1000 & percent.mt <= 25")
```


## nCount_RNA >= 500 & percent.mt <= 10  
```{r}
obj.srt@meta.data %>% filter(predicted_doublet == "False") %>% filter(nCount_RNA >= 500 & percent.mt <= 10) %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("sample", "Number of cells"), caption = "nCount_RNA >= 500 & percent.mt <= 10")
```



## nCount_RNA >= 1000 & percent.mt <= 10  
```{r}
obj.srt@meta.data %>% filter(predicted_doublet == "False") %>% filter(nCount_RNA >= 1000 & percent.mt <= 10) %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(colnames = c("sample", "Number of cells"), caption = "nCount_RNA >= 1000 & percent.mt <= 10")
```

<!-- # UMAP  -->

<!-- ## UMAP by sample -->


<!-- ```{r, fig.width=7, fig.height=6, out.width='80%'} -->
<!-- DimPlot(obj.srt, group.by = 'orig.ident') -->
<!-- ``` -->

<!-- ## UMAP by treatment   -->
<!-- ```{r, fig.width=7, fig.height=6, out.width='80%'} -->
<!-- DimPlot(obj.srt, group.by = 'sample') -->
<!-- ``` -->

<!-- ```{r, fig.width=13, fig.height=4} -->
<!-- DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 4) -->
<!-- ``` -->


<!-- Distribution of number of cells  -->

<!-- ```{r} -->
<!-- obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample') -->
<!-- obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment') -->

<!-- ``` -->


<!-- # Cluster based on the resolutions    -->

<!-- ```{r} -->
<!-- resolution_values <- c(0.2, 0.4, 0.8, 1) -->
<!-- ``` -->

<!-- ## Clusters with the resolution of 0.2   -->
<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=1 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res) -->
<!-- ``` -->

<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=1 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank())  -->
<!-- ``` -->


<!-- Cell number distribution   -->
<!-- ```{r} -->
<!-- # Use !! to unquote the variable within select -->
<!-- #sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다. -->

<!-- # !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다. -->

<!-- obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- ``` -->

<!-- ## Clusters with the resolution of 0.4     -->
<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=2 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res) -->
<!-- ``` -->


<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=2 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank())  -->
<!-- ``` -->


<!-- Cell number distribution   -->
<!-- ```{r} -->

<!-- # Use !! to unquote the variable within select -->
<!-- #sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다. -->

<!-- # !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다. -->

<!-- obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- ``` -->


<!-- ## Clusters with the resolution of 0.8    -->
<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=3 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res) -->
<!-- ``` -->

<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=3 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank())  -->
<!-- ``` -->

<!-- Cell number distribution   -->
<!-- ```{r} -->

<!-- obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- ``` -->


<!-- ## Clusters with the resolution of 1    -->
<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=4 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res) -->
<!-- ``` -->

<!-- ```{r,fig.width=7, fig.height=6, out.width='80%'} -->
<!-- i=4 -->
<!-- res= paste0("RNA_snn_res.", resolution_values[i]) -->
<!-- DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank())  -->
<!-- ``` -->

<!-- ### Cell number distribution   -->
<!-- ```{r} -->
<!-- # Use !! to unquote the variable within select -->
<!-- #sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다. -->

<!-- # !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다. -->

<!-- obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment') -->
<!-- ``` -->


<!-- # Find markers    -->

<!-- resolutions : 0.2, 0.4, 0.8, 1    -->
<!-- ```{r, eval=FALSE, echo=F} -->
<!-- # Marker gene Identification  -->
<!-- find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2,  -->
<!--                                   test_method = 'wilcox', min_percent = 0.25) { -->
<!--   Idents(obj.srt) = cluster_id -->
<!--   all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold),  -->
<!--                                only.pos = TRUE,  -->
<!--                                test.use = test_method, min.pct = min_percent) -->
<!--   return(all.markers) -->
<!-- } -->
<!-- # apply: -->

<!-- resolution_values <- c(0.2, 0.4, 0.8, 1) -->

<!-- # Loop through each resolution value -->
<!-- for (resolution_number in resolution_values) { -->
<!--   # Construct the resolution name -->
<!--   res <- paste0("RNA_snn_res.", resolution_number) -->
<!--   all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res,  -->
<!--                                    logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25) -->
<!--   output_file = paste0(dir,"data/mks/Takeda_", res, ".markers.csv") -->
<!--   write.csv(all.markers, file = output_file, row.names = T) -->
<!-- } -->
<!-- ``` -->


<!-- # Marker genes  -->

<!-- ## resolution 0.2   -->

<!-- Download cluster marker file -->
<!-- ```{r} -->
<!-- resolution_number = 0.2 -->
<!-- res= paste0("RNA_snn_res.", resolution_number) -->
<!-- res -->
<!-- mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1) -->
<!-- len =length(levels(obj.srt@meta.data[,res])) -->
<!-- DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len)) -->
<!-- ``` -->


<!-- ## resolution 0.4   -->

<!-- Download cluster marker file -->
<!-- ```{r} -->
<!-- resolution_number = 0.4 -->
<!-- res= paste0("RNA_snn_res.", resolution_number) -->
<!-- res -->
<!-- mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1) -->
<!-- len =length(levels(obj.srt@meta.data[,res])) -->
<!-- DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len)) -->
<!-- ``` -->


<!-- ## resolution 0.8   -->

<!-- Download cluster marker file -->
<!-- ```{r} -->
<!-- resolution_number = 0.8 -->
<!-- res= paste0("RNA_snn_res.", resolution_number) -->
<!-- res -->
<!-- mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1) -->
<!-- len =length(levels(obj.srt@meta.data[,res])) -->
<!-- DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len)) -->
<!-- ``` -->


<!-- ## resolution 1   -->

<!-- Download cluster marker file -->
<!-- ```{r} -->
<!-- resolution_number = 1 -->
<!-- res= paste0("RNA_snn_res.", resolution_number) -->
<!-- res -->
<!-- mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1) -->
<!-- len =length(levels(obj.srt@meta.data[,res])) -->
<!-- DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len)) -->
<!-- ``` -->
