---
title: "2024 Daiichi DXD"
subtitle: "Initial Report: 30342"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: FALSE
    toc_depth: 6
    toc_float: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align = "left", 
                      message=F, warning=F,
                      results = "markup",
                      error = TRUE,
                      highlight = TRUE,
                      prompt = FALSE,
                      tidy = FALSE)
```

```{r}
## load required packages
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
```

**P30342**


# Pre-processing for 30342    
```{r, eval=FALSE}
# Load necessary libraries
library(Seurat)

# Set base directory and patient information
dir <- '~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD'
patient <- '30342'

# List of folder identifiers
folders <- c('M', 'I', 'A', 'D')

# Initialize an empty list to hold all Seurat objects
obj.list <- list()

# Use a loop to read data from each folder and create Seurat objects
for (folder in folders) {
  # Set the data directory
  dir_features <- paste0(dir, "/raw_data/cellranger_outs/", patient, "/", folder, "/filtered_feature_bc_matrix")
  
  # Read the 10X data
  obj_raw <- Read10X(dir_features)
  
  # Create a Seurat object
  obj_srt <- CreateSeuratObject(counts = obj_raw, project = paste0("P",patient, '_', folder))
  
  # Add the created Seurat object to the list
  obj.list[[folder]] <- obj_srt
}

# Print the list for verification
obj.list

```


```{r}
# Load necessary libraries
library(Seurat)

# Set base directory and patient information
dir <- '~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD'
patient <- '30342'

# List of folder identifiers
folders <- c('M', 'I', 'A', 'D')

# Initialize an empty list to hold all Seurat objects
seurat_objects <- list()

# Use a loop to read data from each folder and create Seurat objects
for (folder in folders) {
  # Set the data directory
  dir_features <- paste0(dir, "/raw_data/cellranger_outs/", patient, "/", folder, "/filtered_feature_bc_matrix")
  
  # Read the 10X data
  obj_raw <- Read10X(dir_features)
  
  # Create a Seurat object
  obj_srt <- CreateSeuratObject(counts = obj_raw, project = paste0("P",patient, '_', folder))
  
  # Rename cells with the folder prefix to distinguish between samples
  obj_srt <- RenameCells(obj_srt, new.names = paste0("P",patient,folder, "_", colnames(obj_srt)))
  
  # Add the created Seurat object to the list
  seurat_objects[[folder]] <- obj_srt
}

# Merge all Seurat objects into a single Seurat object
obj_combined <- Reduce(function(x, y) merge(x, y), seurat_objects)

# Print the combined Seurat object
obj_combined

obj_combined@meta.data[1:3,]
```


```{r}
# Add mitochondrial content 
obj_combined[["percent.mt"]] <- PercentageFeatureSet(obj_combined, pattern = "^MT-")
obj_combined %>% saveRDS(paste0(dir,"/rds/P30342.init.rds"))
```


```{r}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 2000, n_pcs = 20, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.2), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, 
                                  selection.method = 'vst', 
                                  nfeatures = n_features)
  
  # Step 2: Scale and normalize data
  all_genes <- rownames(obj.srt)
  obj.srt <- NormalizeData(obj.srt)
  obj.srt <- ScaleData(obj.srt, features = all_genes)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, 
                    features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj_combined)

# NOTE: SeuratObject 버전 5.0.0에서 GetAssayData 함수가 업데이트되면서 slot 인자가 layer 인자로 변경
```


```{r}
## prepare count mtx using function
write_count_matrix_csv <- function(obj.srt, output_dir, file_name_prefix) {
  data <- obj.srt@assays$RNA@data
  df <- data %>% data.frame() %>% t()
  rownames(df) <- gsub(pattern = '-', replacement = '_', rownames(df))
  output_file <- file.path(output_dir, paste0(file_name_prefix, '.count.mtx.csv'))
  df %>% write.csv(file = output_file)
}

# write_count_matrix_csv
dir <- '~/Desktop/DF/DFCI_Paweletz/2024_Daiichi_DXD/rds/'
output_directory <- dir
file_prefix <- 'P30342'
write_count_matrix_csv(obj.srt=obj.srt, output_directory, file_name_prefix = file_prefix)
```


**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 50000. 

mitochondrial cutoff : 15. 

other cutoffs : NA. 

doublet removal (scrublet) : Y. 

regression: ribosomal genes. 


---

# UMAP 

## UMAP by sample


```{r, fig.width=7, fig.height=6, out.width='80%'}
DimPlot(obj.srt, group.by = 'orig.ident')
```

## UMAP by treatment  
```{r, fig.width=7, fig.height=6, out.width='80%'}
DimPlot(obj.srt, group.by = 'sample')
```

```{r, fig.width=13, fig.height=4}
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 4)
```


Distribution of number of cells 

```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample')
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment')

```


# Cluster based on the resolutions   

```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

## Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}
# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```

## Clusters with the resolution of 0.4    
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```


```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}

# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


## Clusters with the resolution of 0.8   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


## Clusters with the resolution of 1   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

### Cell number distribution  
```{r}
# Use !! to unquote the variable within select
#sym(): 이 함수는 문자열에서 심볼(symbol)을 생성하는 데 사용됩니다. 다시 말해, 문자열을 데이터 프레임이나 환경의 문맥에서 평가할 수 있는 심볼로 변환합니다. 여러분의 경우, 열 이름 (예: "RNA_snn_res.0.2")을 문자열에서 심볼로 변환하는 데 사용됩니다.

# !!: 이것은 비인용 연산자입니다. 이 연산자는 R에게 뒤따라오는 표현식을 tidyverse 함수의 문맥 내에서 평가하도록 지시합니다. !!sym()을 사용하면 사실상 sym()에 의해 생성된 심볼을 평가하고 select() 함수에 대체시킵니다.

obj.srt@meta.data %>% select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Find markers   

resolutions : 0.2, 0.4, 0.8, 1   
```{r, eval=FALSE, echo=F}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)

# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/Takeda_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Marker genes 

## resolution 0.2  

Download cluster marker file
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## resolution 0.4  

Download cluster marker file
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## resolution 0.8  

Download cluster marker file
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## resolution 1  

Download cluster marker file
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```
