---
title: "Takeda NK project"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::readthedown: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

<style> 
#TOC { 
  top: 1%; 
  opacity: 0.5; 
} 
#TOC:hover { 
  opacity: 1; 
} 
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, results = "markup")
options(warn = F)

library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(DT)
```

**Takeda NK**

This is a document for subset data of TN30293 and TN30298.

The details regarding dataset are currently confidential.

<br>
<hr>
<br>

# Import preprocessed data
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_NK/'
obj.srt = readRDS(paste0(dir,'rds/Takeda.NK.30293.30298.subset.23.08.21.rds'))
# obj.srt@meta.data %>% write.csv(paste0(dir,'data/Takeda_NK_filtered.meta.csv'))
```

**pre-processing information**
<br>
<br>
UMI cutoff : 500~ 50000. 

mitochondrial cutoff : 15. 

other cutoffs : NA. 

doublet removal (scrublet) : Y. 

regression: ribosomal genes. 


---

# UMAP 

## UMAP by conditions


### sample
```{r, fig.width=7, fig.height=6, out.width='50%'}
DimPlot(obj.srt, group.by = 'orig.ident')
```
  
### Treatment
```{r, fig.width=7, fig.height=6, out.width='50%'}
DimPlot(obj.srt, group.by = 'sample')
```

```{r, fig.width=12, fig.height=4}
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 4)
```


### Clusters with the resolution of 0.2
```{r,fig.width=7, fig.height=6, out.width='50%'}
DimPlot(obj.srt, group.by = 'RNA_snn_res.0.2')
```

```{r,fig.width=7, fig.height=6, out.width='50%'}
DimPlot(obj.srt, group.by = 'RNA_snn_res.0.2', label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


## The number of cells in each sample 
```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample')
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment')
obj.srt@meta.data %>% select(RNA_snn_res.0.2) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% select(RNA_snn_res.0.2, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(RNA_snn_res.0.2, orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% select(RNA_snn_res.0.2, orig.ident) %>% table() %>% data.frame() %>% tidyr::spread(orig.ident, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and sample')
```

```{r}
df =obj.srt@meta.data %>% select(RNA_snn_res.0.2, orig.ident, sample) %>% table() %>% data.frame()
df$Freq2 = df$Freq
df[df$orig.ident == '30293',]$Freq2 = -(df[df$orig.ident == '30293',]$Freq2)
custom_order <- rev(c(0:13))
df$RNA_snn_res.0.2 = factor(df$RNA_snn_res.0.2, levels = custom_order)
# levels(df$RNA_snn_res.0.2) 
ggplot(df, aes(x = RNA_snn_res.0.2, y = Freq2, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ylim(c(-11000, 11000))
```


```{r}
df= obj.srt@meta.data %>% select(RNA_snn_res.0.2, sample, orig.ident) %>% table() %>% data.frame()
ggplot(df, aes(x = RNA_snn_res.0.2, y = Freq, fill = sample )) +
  geom_bar(stat = "identity", position = "dodge", color="#2C3E50") + facet_wrap(.~orig.ident, ncol = 1) + theme_bw()

```


--- 
# Annotation of cluster to cell type 
```{r, echo=FALSE}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_NK/'
all.markers = read.csv(paste0(dir,'data/mks/30293.30298/Takeda.NK.30293.30298.res.0.2.all.markers.23.08.21.csv'))
## top marker genes table 
mks= all.markers
df=mks$cluster %>% table() %>% data.frame()
df[,'top_genes'] =''
for(i in 1:length(unique(mks$cluster))){
  if(df[i,]$Freq >= 30){
    cl = unique(mks$cluster)[i]
    df[i,'top_genes'] = mks %>% filter(cluster== cl) %>% arrange(desc(avg_log2FC)) %>% head(30) %>% select(gene) %>% pull() %>% paste0(collapse = ',')
  }else{
    cl = unique(mks$cluster)[i]
    df[i,'top_genes'] = mks %>% filter(cluster== cl) %>% arrange(desc(avg_log2FC)) %>% select(gene) %>% pull() %>% paste0(collapse = ',')
  }
}
```
  
## Top 30 genes table
```{r}
DT::datatable(df, caption = 'Top 30 genes') 
```



# cell type marker 

## Dendritic cells 
CD83: CD83 is a marker associated with mature dendritic cells.  
CD86: CD86 is an activation marker that is upregulated on dendritic cells upon activation.    
CD123 (IL3RA): CD123 is a marker for plasmacytoid dendritic cells (pDCs), which produce type I interferons in response to viral infections.  
HLA-DR: Human Leukocyte Antigen-DR (HLA-DR) is an MHC class II molecule expressed on the surface of dendritic cells involved in antigen presentation.  
CD80 (B7-1) and CD86 (B7-2): CD80 and CD86 are co-stimulatory molecules expressed on dendritic cells upon activation, promoting T cell activation.  
CD1c (BDCA-1): CD1c is another marker used to identify myeloid dendritic cells.  


# Cell type guess by common signature genes  

## Cell type- signatures  
```{r}
# Create a list of categories and associated genes
categories <- list(
  B = c("CD79A", "IGHM", "CD79B", "MS4A1"),
  Macrophage = c("CD14", "VSIG4", "C1QB", "C1QA", "APOE", "SPP1", "CD83"),
  DC = c("ITGAX", "ITGAM", "CD1C", "FLT3", "HLA-DRA", "HLA-DRB1","CD83","CD86","IL3RA"),
  Fibroblast = c("COL1A1", "COL1A2", "COL3A1", "COL4A1", "COL4A2", "S100A16", "UGDH"),
  Tcell = c("TCF7", "CD3G", "CD3E", "CD3D", "CD4", "CD28", "BCL11B", "CD8A", "CD8B", "ICOS"),
  NK = c("GNLY", "KLRD1", "KLRB1", "KLRC1"),
  RBC = c("HBB"),
  Neutrophils = c("CEACAM8","FUT4","CD63"), 
  Epithelial = c("KRT18", "KRT8", "KRT18", "KRT19"),
  Tumor = c("CTLA4", "HAVCR2", "EGFR", "KRAS", "MYC", "CD274", "EPCAM"),
  Monocyte = c("CD14", "CD68", "CCR2", "CSF1R")
)

# Print the list
print(categories)
```

## Read marker genes  
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_NK/'
all.markers = read.csv(paste0(dir,'data/mks/30293.30298/Takeda.NK.30293.30298.res.0.2.all.markers.23.08.21.csv'), row.names = 1)
all.markers %>% head()
```

# cluster summary for cell types. 
```{r}
generate_cluster_summary <- function(cl, percent_in_cluster, percent_out_cluster) {
  # Compute cluster-specific genes
  cluster_specific_genes <- all.markers %>%
    filter(cluster == cl) %>%
    filter(pct.1 >= percent_in_cluster & pct.2 <= percent_out_cluster) %>%
    top_n(50, avg_log2FC) %>%
    dplyr::select(gene) %>%
    pull()
  
  # Create a data frame for summary
  df <- data.frame(
    cluster = cl,
    category = names(categories),
    stringsAsFactors = FALSE
  )
  
  # Create a function to perform the intersection and handle "NA"
  intersect_with_na <- function(genes, category_genes) {
    result <- intersect(genes, category_genes)
    if (length(result) == 0) {
      return("no matching genes")
    } else {
      return(paste(result, collapse = ", "))
    }
  }
  
  # Apply the intersection function and compute the number of matches
  df$genes <- sapply(1:nrow(df), function(j) {
    intersect_with_na(cluster_specific_genes, categories[[j]])
  })
  
  # If needed, you can convert the result to a character vector
  df$genes <- as.character(df$genes)
  
  # Return the summary data frame
  return(df)
}

cluster_specific_genes = function(cl, percent_in_cluster, percent_out_cluster) {
  # Compute cluster-specific genes
  cluster_specific_genes <- all.markers %>%
    filter(cluster == cl) %>%
    filter(pct.1 >= percent_in_cluster & pct.2 <= percent_out_cluster) %>%
    top_n(50, avg_log2FC) %>%
    dplyr::select(gene) %>%
    pull()
  return(cluster_specific_genes)
}

cluster_defining_genes = function(cl) {
  # Compute cluster-specific genes
  cluster_specific_genes <- all.markers %>%
    filter(cluster == cl) %>%
    
    top_n(50, avg_log2FC) %>%
    dplyr::select(gene) %>%
    pull()
  return(cluster_specific_genes)
}
```
   
**For excluesivly-expressed gene filtering **   
percent_in_cluster = 0.6.  
percent_out_cluster = 0.5.  

## cluster 0 :
```{r}
i <- 0
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 1 :
```{r}
i <- 1
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 2 :
```{r}
i <- 2
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 3 :
```{r}
i <- 3
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 4 :
```{r}
i <- 4
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 5 :
```{r}
i <- 5
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 6 :
```{r}
i <- 6
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 7 :
```{r}
i <- 7
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 8 :
```{r}
i <- 8
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 9 :
```{r}
i <- 9
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 10 :
```{r}
i <- 10
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 11 :
```{r}
i <- 11
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 12 :
```{r}
i <- 12
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```


## cluster 13 :
```{r}
i <- 13
percent_in_cluster <- 0.6
percent_out_cluster <- 0.5
result_df <- generate_cluster_summary(i, percent_in_cluster, percent_out_cluster)
cluster_specific_genes_output= cluster_specific_genes(i, percent_in_cluster, percent_out_cluster)
cluster_defining_genes_output= cluster_defining_genes(i)
cat("cluster specific genes" ,"\n")
cluster_specific_genes_output
cat("cluster defininig genes" ,"\n")
cluster_defining_genes_output 
result_df 
```

