---
title: "Takeda PD1 + NK set merged"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Merging condition  

4 from Elena's set    
MT30273.raw.23.05.16.rds    
MT30271.raw.23.05.16.rds    
MT30269.raw.23.05.16.rds    
MT30212.raw.23.05.16.rds    
    
2 from Patrick's set    
Takeda.NK.30298.23.08.17.rds    
Takeda.NK.30293.23.08.17.rds    
    
    


```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>

# Data Overview     

  
```{r, eval=F}
dir='~/Desktop/DF/DFCI_Paweletz/'

obj1= readRDS(paste0(dir,"2023_Takeda/rds/MT30273.raw.23.05.16.rds"))
obj2= readRDS(paste0(dir,"2023_Takeda/rds/MT30271.raw.23.05.16.rds"))
obj3= readRDS(paste0(dir,"2023_Takeda/rds/MT30269.raw.23.05.16.rds"))
obj4= readRDS(paste0(dir,"2023_Takeda/rds/MT30212.raw.23.05.16.rds"))
obj5= readRDS(paste0(dir,"2023_Takeda_NK/rds/Takeda.NK.30298.23.08.17.rds"))
obj6= readRDS(paste0(dir,"2023_Takeda_NK/rds/Takeda.NK.30293.23.08.17.rds"))

# ## merge 
obj.srt = merge(obj1, y = c(obj2,obj3,obj4,obj5,obj6), 
                add.cell.ids = c('MT30273','MT30271','MT30269',"MT30212", "NK30298","NK30293"))

```

Total 126229 cells from 6 samples 


| orig.ident |  COM  | CTL  | NKC  | PEM  | TAK  |
|------------|------|------|------|------|------|
|   30293    | 4176 | 5058 | 7463 | 0    | 6472 |
|   30298    | 8578 | 11398| 7131 | 0    | 7595 |
|  MT30212   | 867  | 883  | 0    | 1005 | 656  |
|  MT30269   | 4553 | 7118 | 0    | 10693| 6287 |
|  MT30271   | 3646 | 6091 | 0    | 7853 | 6835 |
|  MT30273   | 5173 | 3929 | 0    | 0    | 2769 |


```{r, eval=FALSE}
obj.srt = subset(obj.srt, sample %in% c("CTL","TAK"))
#obj.srt@meta.data[,c('orig.ident','sample')] %>% table() 
obj.srt <- subset(obj.srt, predicted_doublet == 'False' & nCount_RNA >= 500 & nCount_RNA < 20000 & percent.mt <= 25)
#obj.srt@meta.data[,c('orig.ident','sample')] %>% table()

```


Selected samples    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 5058 | 6472 |
|   30298    | 11398| 7595 |
|  MT30212   | 883  | 656  |
|  MT30269   | 7118 | 6287 |
|  MT30271   | 6091 | 6835 |
|  MT30273   | 3929 | 2769 |



Selected samples (after filtering)    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 3976 | 5182 |
|   30298    | 9620 | 6054 |
|  MT30212   | 522  | 463  |
|  MT30269   | 6185 | 5551 |
|  MT30271   | 5602 | 6412 |
|  MT30273   | 3031 | 1569 |


Filtering  
- doublet removal   
- UMI : 500 < & < 20,000    
- mt content <= 25    
- ribosomal gene regression out    


```{r, eval=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  ## Check ribosomal gene ratio and add ribosomal gene percentage
  ribosomal_genes <- rownames(obj.srt)[grepl('^RP', rownames(obj.srt))]
  obj.srt[["percent.ribo"]] <- PercentageFeatureSet(obj.srt, pattern = "^RP")
  
  ## Define variables in metadata to regress
  vars_to_regress <- c('percent.ribo')
  
  # Step 2: Scale and normalize data
  ## Regress out the uninteresting sources of variation in the data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(object = obj.srt, vars.to.regress = vars_to_regress, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
obj.srt %>% saveRDS(paste0(dir,'rds/Takeda_6conditions.23.12.01.rds'))
```


# Data   
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
obj.srt =readRDS(paste0(dir,'rds/Takeda_6conditions.23.12.01.rds'))
#obj.srt %>% saveRDS(paste0(dir,'rds/Takeda_6conditions.23.12.01.rds'))
```


Total number of cells: 54167    


```{r,eval=FALSE}
# Total number of cells 
obj.srt@meta.data %>% nrow()
```


## UMAP    

### Distribution of samples   
```{r, fig.height=7, fig.width=8}
DimPlot(obj.srt, group.by = "orig.ident", pt.size = 0.1)
```

```{r, fig.height=7, fig.width=12}
DimPlot(obj.srt, group.by = "orig.ident", pt.size = 0.1, split.by = "orig.ident", ncol = 3)
```

### Distribution of Treatment   
```{r, fig.height=7, fig.width=8}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1)
```

```{r, fig.height=7, fig.width=12}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1, split.by = "sample", ncol = 3)
```


### Distribution of samples splited by treatment   


```{r, fig.height=12, fig.width=8}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1, split.by = "orig.ident", ncol = 2)
```


## Clusters   

```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

### Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


### Clusters with the resolution of 0.4  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


### Clusters with the resolution of 0.8  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

### Clusters with the resolution of 1  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

# Composition of each cluster based on % of cells from each sample   

## res 0.2  
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 1  
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```


# Marker genes based on clusters    

```{r, eval=FALSE, echo=F}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
# Find Markers based on cluster
# Define function
# Apply function
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)
# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/Merge6_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```



## 0.2  

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
```


Download cluster marker file
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```




## 0.4  

Download cluster marker file
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```




## 0.8  

Download cluster marker file
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```




## 1  

Download cluster marker file
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


