---
title: "Takeda Meloid from Merged dataset"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Report 

This is a report for meyloid cell subset from merged dataset.

```{r, echo=F}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>

# Import data   


```{r}
celltypes_to_subset = c("lymphoid", "mesenchymal", "myeloid")
cat("Three cell types are ", "\n")
cat(celltypes_to_subset, "\n") 
cat("\n")
cat("The dataset here is ", celltypes_to_subset[3], "\n")

```


```{r, eval=FALSE}
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda/"
obj.srt= readRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[3], ".rds")))
```

```{r, eval=FALSE}
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  ## Check ribosomal gene ratio and add ribosomal gene percentage
  ribosomal_genes <- rownames(obj.srt)[grepl('^RP', rownames(obj.srt))]
  obj.srt[["percent.ribo"]] <- PercentageFeatureSet(obj.srt, pattern = "^RP")
  
  ## Define variables in metadata to regress
  vars_to_regress <- c('percent.ribo')
  
  # Step 2: Scale and normalize data
  ## Regress out the uninteresting sources of variation in the data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(object = obj.srt, vars.to.regress = vars_to_regress, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[3], ".rds")))
```

<br>

<br>


**pre-processing information** 

UMI cutoff : 500~ 20000.

mitochondrial cutoff : 25.

other cutoffs : NA.

doublet removal (scrublet) : Y.

regression: ribosomal genes.


# Import processed data    

MT_merged.3_conditions_mesenchymal cells        
```{r}
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda/"
obj.srt = readRDS(file.path(dir, paste0("rds/obj.MT_merged.3_conditions.", celltypes_to_subset[3], ".rds")))
```

```{r}
obj.srt@meta.data$sample = factor(obj.srt@meta.data$sample, levels = c("CTL","TAK","COM"))
```


# UMAP 

## UMAP by sample
   
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'orig.ident')
```


```{r, fig.width=14, fig.height=5}
DimPlot(obj.srt, group.by = 'orig.ident', split.by = 'orig.ident', ncol = 3)
```

## UMAP by treatment    
```{r, fig.width=7, fig.height=6, out.width='70%'}
DimPlot(obj.srt, group.by = 'sample')
```

```{r, fig.width=14, fig.height=5}
DimPlot(obj.srt, group.by = 'sample', split.by = 'sample', ncol = 3)
```

## Distribution of number of cells 

```{r}
obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample')
obj.srt@meta.data %>% select(orig.ident,sample) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by sample and treatment')

```


```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

# Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}

obj.srt@meta.data %>% dplyr::select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% dplyr::select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% dplyr::select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```

# Clusters with the resolution of 0.4   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```


```{r,fig.width=7, fig.height=6, out.width='80%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


Cell number distribution  
```{r}
obj.srt@meta.data %>% dplyr::select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% dplyr::select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% dplyr::select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 0.8   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}
obj.srt@meta.data %>% dplyr::select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% dplyr::select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% dplyr::select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```


# Clusters with the resolution of 1   
```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='80%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

Cell number distribution  
```{r}

obj.srt@meta.data %>% dplyr::select(!!sym(res)) %>% table() %>% data.frame() %>% DT::datatable(caption = 'Number of Cells by clusters')
obj.srt@meta.data %>% dplyr::select(!!sym(res), sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
obj.srt@meta.data %>% dplyr::select(!!sym(res), orig.ident, sample) %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable(caption = 'Number of Cells by clusters and treatment')
```



**Find markers**   

resolution : 0.2, 0.4, 0.8, 1   
```{r, eval=FALSE, echo=F}
# Marker gene Identification 
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda/'
# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/Takeda_", celltypes_to_subset[3],"_",res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


# Marker genes 

## 0.2  

Download cluster marker file   
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[3],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.4  

Download cluster marker file   
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[1],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 0.8  

Download cluster marker file   
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[3],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


## 1  

Download cluster marker file   
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
mks = read.csv(paste0(dir,"data/mks/Takeda_", celltypes_to_subset[3],"_",res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```



# Annotation of clusters  

## 0.2   
```{r}
library(DT)

resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.4   
```{r}
library(DT)

resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 0.8     
```{r}
library(DT)

resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


## 1   
```{r}
library(DT)

resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
n=length(levels(obj.srt@meta.data[, res]))
df = data.frame(cluster =paste0("cluster", levels(obj.srt@meta.data[, res])),
                Proposed_cell_type = rep("NA", n),
                Cell_type_markers = rep("NA", n))
DT::datatable(df, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=n))
```


Kindly complete the cluster annotation that you want to dplyr::select and send it to Sung for update.  



# Composition of each cluster based on % of cells from each sample   

One cluster : consist of 3 samples (% of sample1, % of sample2, % of sample3)  

## res 0.2 
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```


## res 0.4  
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```



```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100)%>% DT::datatable()
```


## res 0.8  
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=12, fig.height=6}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```



## res 1 
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
res
```

```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') +theme_minimal()
```


```{r, fig.width=14, fig.height=8}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ggplot(aes(x = !!sym(res), y = percent, fill = orig.ident)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.1f%%", percent)), 
            position = position_stack(vjust = 0.5),
            show.legend = FALSE, size=3) +
  labs(y = "Percentage") +
  theme_minimal()
```

```{r}
obj.srt@meta.data %>%
  group_by(!!sym(res), orig.ident) %>%
  summarise(count = n()) %>%
  group_by(!!sym(res)) %>%
  mutate(percent = count / sum(count) * 100) %>% DT::datatable()
```






