---
title: "Takeda PD1 + NK set merged"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Merging condition  

4 from Elena's set    
MT30273.raw.23.05.16.rds    
MT30271.raw.23.05.16.rds    
MT30269.raw.23.05.16.rds    
MT30212.raw.23.05.16.rds    
    
2 from Patrick's set    
Takeda.NK.30298.23.08.17.rds    
Takeda.NK.30293.23.08.17.rds    
    
    


```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>


Selected samples (after filtering)    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 3976 | 5182 |
|   30298    | 9620 | 6054 |
|  MT30212   | 522  | 463  |
|  MT30269   | 6185 | 5551 |
|  MT30271   | 5602 | 6412 |
|  MT30273   | 3031 | 1569 |


Filtering  
- doublet removal   
- UMI : 500 < & < 20,000    
- mt content <= 25    
- ribosomal gene regression out    


# Data   
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
obj.srt =readRDS(paste0(dir,'rds/Takeda_6conditions.23.12.01.rds'))
```


Total number of cells: 54167    


```{r,eval=FALSE}
# Total number of cells 
obj.srt@meta.data %>% nrow()
```


# Gene expression   


## geneset  
```{r}
df <- data.frame(
  Category = c("B" ,"Fibroblast","Myeloid","T_NK","Tumor"),
  Genes = c("CD79A, IGHM, CD79B",
            "COL3A1, COL4A1, COL4A2, PECAM1",
            "CD14, VSIG4, C1QB, C1QA, APOE, SPP1, CD83, ITGAX, ITGAM, CD1C,FLT3, HLA-DRA, HLA-DRB1, CD68, CCR2, CSF1R",
            "TCF7, CD3G, CD3E, CD3D, CD4, CD28, CD8A, CD8B, ICOS, GNLY, KLRD1, KLRB1, KLRC1",
            "KRT18, KRT8, KRT18, KRT19, EGFR, KRAS, MYC, EPCAM, MSLN, PDPN"
            )
)

gene_list <- lapply(strsplit(df$Genes, ", "), as.vector)
names(gene_list) <- df$Category

# Print the resulting list

df %>% as.data.frame() %>% gt::gt()
```

```{r}
gene_list <- strsplit(df$Genes, ", ")
names(gene_list) <- df$Category 
```


## B   

```{r, fig.width=8, fig.height=7}
FeaturePlot(obj.srt, features = gene_list$B, pt.size = 0.1, ncol = 2)
```


## Fibroblast   

```{r, fig.width=8, fig.height=7}
FeaturePlot(obj.srt, features = gene_list$Fibroblast, pt.size = 0.1, ncol = 2)
```

## Myeloid   

```{r, fig.width=13, fig.height=7}
FeaturePlot(obj.srt, features = gene_list$Myeloid, pt.size = 0.1, ncol = 5)
```


## T_NK   

```{r, fig.width=10, fig.height=8}
FeaturePlot(obj.srt, features = gene_list$T_NK, pt.size = 0.1, ncol = 4)
```


## Tumor
```{r, fig.width=10, fig.height=6}
FeaturePlot(obj.srt, features = gene_list$Tumor, pt.size = 0.1, ncol = 4)
```


# Gene expression in resolution 0.2      
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
Idents(obj.srt) = res
```


## B   
```{r}
VlnPlot(obj.srt, features = gene_list$B, stack = T, flip = T)
```

## Fibroblast   
```{r}
VlnPlot(obj.srt, features = gene_list$Fibroblast, stack = T, flip = T)
```


## Myeloid   
```{r}
VlnPlot(obj.srt, features = gene_list$Myeloid, stack = T, flip = T)
```

## T_NK   
```{r}
VlnPlot(obj.srt, features = gene_list$T_NK, stack = T, flip = T)
```

## Tumor   
```{r}
VlnPlot(obj.srt, features = gene_list$Tumor, stack = T, flip = T)
```


# Gene expression in resolution 0.4      
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
Idents(obj.srt) = res
```


## B   
```{r}
VlnPlot(obj.srt, features = gene_list$B, stack = T, flip = T)
```

## Fibroblast   
```{r}
VlnPlot(obj.srt, features = gene_list$Fibroblast, stack = T, flip = T)
```

## Myeloid   
```{r}
VlnPlot(obj.srt, features = gene_list$Myeloid, stack = T, flip = T)
```

## T_NK   
```{r}
VlnPlot(obj.srt, features = gene_list$T_NK, stack = T, flip = T)
```

## Tumor   
```{r}
VlnPlot(obj.srt, features = gene_list$Tumor, stack = T, flip = T)
```



# Gene expression in resolution 0.8      
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
Idents(obj.srt) = res
```


## B   
```{r}
VlnPlot(obj.srt, features = gene_list$B, stack = T, flip = T)
```

## Fibroblast   
```{r}
VlnPlot(obj.srt, features = gene_list$Fibroblast, stack = T, flip = T)
```

## Myeloid   
```{r}
VlnPlot(obj.srt, features = gene_list$Myeloid, stack = T, flip = T)
```

## T_NK   
```{r}
VlnPlot(obj.srt, features = gene_list$T_NK, stack = T, flip = T)
```

## Tumor   
```{r}
VlnPlot(obj.srt, features = gene_list$Tumor, stack = T, flip = T)
```



# Gene expression in resolution 1         
```{r}
resolution_number =1
res= paste0("RNA_snn_res.", resolution_number)
Idents(obj.srt) = res
```


## B   
```{r}
VlnPlot(obj.srt, features = gene_list$B, stack = T, flip = T)
```

## Fibroblast   
```{r}
VlnPlot(obj.srt, features = gene_list$Fibroblast, stack = T, flip = T)
```

## Myeloid   
```{r}
VlnPlot(obj.srt, features = gene_list$Myeloid, stack = T, flip = T)
```

## T_NK   
```{r}
VlnPlot(obj.srt, features = gene_list$T_NK, stack = T, flip = T)
```

## Tumor   
```{r}
VlnPlot(obj.srt, features = gene_list$Tumor, stack = T, flip = T)
```