---
title: "Takeda PD1 + NK merged: Lymphoid"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)
```



```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>

# Data Overview     

Total 126229 cells from 6 samples 


| orig.ident |  COM  | CTL  | NKC  | PEM  | TAK  |
|------------|------|------|------|------|------|
|   30293    | 4176 | 5058 | 7463 | 0    | 6472 |
|   30298    | 8578 | 11398| 7131 | 0    | 7595 |
|  MT30212   | 867  | 883  | 0    | 1005 | 656  |
|  MT30269   | 4553 | 7118 | 0    | 10693| 6287 |
|  MT30271   | 3646 | 6091 | 0    | 7853 | 6835 |
|  MT30273   | 5173 | 3929 | 0    | 0    | 2769 |


Selected samples    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 5058 | 6472 |
|   30298    | 11398| 7595 |
|  MT30212   | 883  | 656  |
|  MT30269   | 7118 | 6287 |
|  MT30271   | 6091 | 6835 |
|  MT30273   | 3929 | 2769 |



Selected samples (after filtering)    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 3976 | 5182 |
|   30298    | 9620 | 6054 |
|  MT30212   | 522  | 463  |
|  MT30269   | 6185 | 5551 |
|  MT30271   | 5602 | 6412 |
|  MT30273   | 3031 | 1569 |


Filtering  
- doublet removal   
- UMI : 500 < & < 20,000    
- mt content <= 25    
- ribosomal gene regression out    


# Lymphoid Data   

```{r}
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
obj.srt = readRDS(paste0(dir,"rds/obj.cond6_merged.lymphoid.rds"))
```

## Number of Lymphoid cells  
```{r}
df1 =obj.srt@meta.data %>% select(orig.ident) %>% table() %>% data.frame()
df2 =obj.srt@meta.data %>% select(orig.ident, sample) %>% table() %>% 
  data.frame() %>% tidyr::spread(sample, Freq)
df = cbind(df1, df2[,2:3])
colnames(df)[1] = "Total"
df %>% DT::datatable()
```

```{r, eval=FALSE}
file_name= "obj.cond6_merged.lymphoid.rds"
## perform default analysis
perform_default_analysis <- function(obj.srt, n_features = 3000, n_pcs = 30, 
                                     dims_for_neighbors = 1:20, 
                                     resolutions = c(0.1, 0.2, 0.4, 0.8,1), 
                                     umap_dims = 1:20) {
  # Step 1: Find variable features
  obj.srt <- FindVariableFeatures(obj.srt, selection.method = 'vst', nfeatures = n_features)
  
  ## Check ribosomal gene ratio and add ribosomal gene percentage
  ribosomal_genes <- rownames(obj.srt)[grepl('^RP', rownames(obj.srt))]
  obj.srt[["percent.ribo"]] <- PercentageFeatureSet(obj.srt, pattern = "^RP")
  
  ## Define variables in metadata to regress
  vars_to_regress <- c('percent.ribo')
  
  # Step 2: Scale and normalize data
  ## Regress out the uninteresting sources of variation in the data
  all_genes <- rownames(obj.srt)
  obj.srt <- ScaleData(object = obj.srt, vars.to.regress = vars_to_regress, features = all_genes)
  obj.srt <- NormalizeData(obj.srt)
  
  # Step 3: Run PCA
  obj.srt <- RunPCA(obj.srt, features = VariableFeatures(object = obj.srt), npcs = n_pcs)
  
  # Step 4: Find neighbors
  obj.srt <- FindNeighbors(obj.srt, dims = dims_for_neighbors)
  
  # Step 5: Find clusters
  obj.srt <- FindClusters(obj.srt, resolution = resolutions)
  
  # Step 6: Run UMAP
  obj.srt <- RunUMAP(obj.srt, dims = umap_dims)
  
  # Return the Seurat object with analysis results
  return(obj.srt)
}

# apply
obj.srt <- perform_default_analysis(obj.srt)
obj.srt %>% saveRDS(paste0(dir,'rds/',file_name))
```


## UMAP    

### Distribution of samples   
```{r, fig.height=7, fig.width=8, out.width="60%"}
DimPlot(obj.srt, group.by = "orig.ident", pt.size = 0.1)
```

```{r, fig.height=7, fig.width=12}
DimPlot(obj.srt, group.by = "orig.ident", pt.size = 0.1, split.by = "orig.ident", ncol = 3)
```

### Distribution of Treatment   
```{r, fig.height=7, fig.width=8, out.width="60%"}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1)
```

```{r, fig.height=7, fig.width=12, out.width="60%"}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1, split.by = "sample", ncol = 3)
```


### Distribution of samples splited by treatment   


```{r, fig.height=7, fig.width=8, out.width="80%"}
DimPlot(obj.srt, group.by = "sample", pt.size = 0.1, split.by = "orig.ident", ncol = 2)
```

## Clusters   

```{r}
resolution_values <- c(0.2, 0.4, 0.8, 1)
```

### Clusters with the resolution of 0.2  
```{r,fig.width=7, fig.height=6, out.width='60%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='60%'}
i=1
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


### Clusters with the resolution of 0.4  
```{r,fig.width=7, fig.height=6, out.width='60%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='60%'}
i=2
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


### Clusters with the resolution of 0.8  
```{r,fig.width=7, fig.height=6, out.width='60%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='60%'}
i=3
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```

### Clusters with the resolution of 1  
```{r,fig.width=7, fig.height=6, out.width='60%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res)
```

```{r,fig.width=7, fig.height=6, out.width='60%'}
i=4
res= paste0("RNA_snn_res.", resolution_values[i])
DimPlot(obj.srt, group.by = res, label = T, label.box = T, label.size = 3) + theme(plot.title = element_blank()) 
```


# Composition of each cluster based on % of cells from each sample   

## res 0.2  
```{r, out.width="80%"}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 0.4  
```{r, out.width="80%"}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 0.8  
```{r, out.width="80%"}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```

## res 1  
```{r, out.width="80%"}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
obj.srt@meta.data %>% ggplot(aes(!!sym(res), fill=orig.ident)) + geom_bar(position = 'fill') + theme_minimal()
```


# Marker genes based on clusters    
```{r, eval=FALSE, echo=F}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
# Find Markers based on cluster
# Define function
# Apply function
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  return(all.markers)
}
# apply:

resolution_values <- c(0.2, 0.4, 0.8, 1)
# Loop through each resolution value
for (resolution_number in resolution_values) {
  # Construct the resolution name
  res <- paste0("RNA_snn_res.", resolution_number)
  all.markers= find_and_save_markers(obj.srt= obj.srt, cluster_id = res, 
                                   logfc_threshold = 1.2, test_method = 'wilcox', min_percent = 0.25)
  output_file = paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv")
  write.csv(all.markers, file = output_file, row.names = T)
}
```


## 0.2  

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
```


Download cluster marker file
```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```

## 0.4  

Download cluster marker file
```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```




## 0.8  

Download cluster marker file
```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
len =length(levels(obj.srt@meta.data[,res]))
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```




## 1  

Download cluster marker file
```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
DT::datatable(mks, editable = TRUE, extensions = "Buttons", options = list(dom="Bfrtip", buttons=c("csv","excel"), pageLength=len))
```


# Heatmaps    


## 0.2  

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
```


```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=12, fig.height=12}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 0.4  

```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=14, fig.height=14}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 0.8  

```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=16, fig.height=20}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 1  


```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_lymphoid_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=22, fig.height=28}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 10))
```

<br><br>

# CD8A+ population in dataset   


```{r}
genes = c("CD8A")
FeaturePlot(obj.srt, features = genes)
```


```{r}
# update gene information to meta data
update_gene_Status <- function(obj, g) {
  exp <- obj@assays$RNA@data[g, ]
  
  # Assuming that 'CD8A' is already a column in 'obj@meta.data'
  obj@meta.data[, g] <- exp
  
  # Create a new column 'CD8A_pos' based on the expression level of CD8A
  obj@meta.data <- obj@meta.data %>%
    mutate(!!paste0(g, "_pos") := ifelse(.data[[g]] > 0, paste(g, "+"), paste(g, "-")))
  
  return(obj)
}
# Example usage:
obj.srt <- update_gene_Status(obj.srt, 'CD8A')
```




```{r}
## CD8A+ population in UMAP   
#DimPlot(obj.srt, group.by = "CD8A_pos", cols = c("grey","red"))
```


```{r, eval=FALSE}
DimPlot(obj.srt, group.by = "CD8A_pos", cols = c("grey","red"), 
        split.by = 'orig.ident', ncol = 3)
```

```{r,fig.width=8, fig.height=8}
obj.srt@meta.data$sample_treatment = paste0(obj.srt@meta.data$orig.ident,"_", obj.srt@meta.data$sample)
# DimPlot(obj.srt, group.by = "CD8A_pos", cols = c("grey","red"), 
#         split.by = 'sample_treatment', ncol = 4)
```



```{r}
## CD8A+ population ratio    
group = "sample_treatment"
gene_variable = "CD8A_pos"
# obj.srt@meta.data %>% ggplot(aes(!!sym(group), fill=!!sym(gene_variable))) + 
#   geom_bar(position = 'fill') + theme_minimal() + 
#   scale_fill_manual(values = c("#F4F6F6","#EC7063")) +
#   RotatedAxis()
```


```{r}
## Effector T cell related gene expression in CD8A+ population   

# Assuming 'CD8A_pos' is a column in obj.srt@meta.data
CD8A_positive_cells <- obj.srt@meta.data %>%
  filter(CD8A_pos == "CD8A +")

# Subset obj.srt to include only CD8A+ cells
obj_srt_CD8A_positive <- subset(obj.srt, cells = rownames(CD8A_positive_cells))

# Create the violin plot
# VlnPlot(obj_srt_CD8A_positive, features = c("GZMA", "GZMB","GZMH","NKG7"), 
#         group.by = "sample_treatment", stack = TRUE, flip = TRUE)

```


## IFNG-related gene expression in CD8A+ population   
```{r, fig.width=8, fig.height=8}
# Create the violin plot
VlnPlot(obj_srt_CD8A_positive, features = c("IFNG","IFI44","IFIT2","IFIT3","ISG15", "ISG20"), 
        group.by = "sample_treatment", stack = TRUE, flip = TRUE)

```



```{r, eval=FALSE}
genes = c("CD79A", "CD4","CD8A","NCAM1","NCAM2","FCGR3A","FCGR3B", "CD3E","CD3D")
FeaturePlot(obj.srt, features = genes, pt.size = 0.1, ncol = 3)

```


```{r, fig.width=8, fig.height=8, eval=FALSE}
## Tem, Trm related gene expression in CD8A+ population   

genes=c("CCR7","SELL","PTPRC","CCR5","CD27","CD44","IL7R","EOMES",
        "CD69","ITGAE","CXCR6","CD4")
# Create the violin plot
VlnPlot(obj_srt_CD8A_positive, features = genes, 
        group.by = "sample_treatment", stack = TRUE, flip = TRUE)
# CD69 is increased by TAK treatment.   
# 
# (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10010323/)    

```



## Requested   
```{r, fig.width=8, fig.height=12}
# New order for plotting
obj_srt_CD8A_positive@meta.data$grp_order = factor(obj_srt_CD8A_positive@meta.data$sample_treatment)
levels(obj_srt_CD8A_positive@meta.data$grp_order) =c("MT30212_CTL", "MT30212_TAK",
                                                     "MT30271_CTL", "MT30271_TAK",
                                                     "MT30269_CTL", "MT30269_TAK",
                                                     "MT30273_CTL", "MT30273_TAK",
                                                     "30293_CTL", "30293_TAK",
                                                     "30298_CTL","30298_TAK")
genes <- c("TIGIT", "LAG3", "ATM", "COX6B1", "PNRC1", "TUBB", "ACTB", "GAPDH")
# Create the violin plot
VlnPlot(obj_srt_CD8A_positive, features = genes, 
        group.by = "grp_order", stack = TRUE, flip = TRUE)
```


```{r, fig.width=8, fig.height=16}
ids = c("MT30212_CTL","MT30271_CTL","MT30269_CTL","MT30273_CTL","30293_CTL","30298_CTL")
Idents(obj_srt_CD8A_positive) = "sample_treatment"
obj.sub = subset(obj_srt_CD8A_positive, idents = ids)
genes <- c("PTPRC", "CD3E", "CD3D", "CD8B", "CD2", "FYN", "CD28", "ITGAL", "RABGAP1L", "GBP4", "GBP2", "SOD2", "STK17A", "CD84", "CD99")
VlnPlot(obj.sub, features = genes, 
        group.by = "grp_order", stack = TRUE, flip = TRUE)
```



```{r, eval=FALSE}
# Comparison in CD8+T population   

## TAK / CTL   
obj = obj_srt_CD8A_positive
Idents(obj) = 'sample'
logfc=log2(1)
mks =FindMarkers(obj, ident.1 = 'TAK', ident.2 = 'CTL', 
                 logfc.threshold = logfc)
pval=0.05
fc=1.2
mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                               ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))

mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))

mks = mks %>% mutate(color = ifelse(is.na(p_val_adj), "no_sig",
                                    ifelse(log10(p_val_adj) <= log10(0.05) & 
                                             avg_log2FC >= log2(1.2), "UP", 
                                           ifelse(log10(p_val_adj) <= log10(0.05) & 
                                                    avg_log2FC <= -log2(1.2), "DN", "no_sig"))))
mks$color = factor(mks$color, levels = c("UP","DN","no_sig"))

# Volcanoplot 
mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color=color)) + 
  geom_point(size=1, alpha=0.5) + theme_classic() + 
  ggtitle("TAK / CTL")+
  scale_color_manual(values = c("red","blue",'grey')) +
  geom_vline(xintercept = c(-log2(1.2), log2(1.2)), color='red') +
  geom_hline(yintercept = -log10(0.05), color='blue') + theme_classic() 
mks.all = mks
```



```{r, eval=FALSE}
## R / NR in all   
obj_srt_CD8A_positive$response = factor(obj_srt_CD8A_positive$orig.ident)
levels(obj_srt_CD8A_positive$response) = c("NR","NR","R","NR","R","NR")

obj = obj_srt_CD8A_positive

Idents(obj) = 'response'
logfc=log2(1)
mks =FindMarkers(obj, ident.1 = 'R', ident.2 = 'NR', 
                 logfc.threshold = logfc, test.use="DESeq2")
pval=0.05
fc=1.2
mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                               ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))

mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))

mks = mks %>% mutate(color = ifelse(is.na(p_val_adj), "no_sig",
                                    ifelse(log10(p_val_adj) <= log10(0.05) & 
                                             avg_log2FC >= log2(1.2), "UP", 
                                           ifelse(log10(p_val_adj) <= log10(0.05) & 
                                                    avg_log2FC <= -log2(1.2), "DN", "no_sig"))))
mks$color = factor(mks$color, levels = c("UP","DN","no_sig"))
mks.rnr = mks
# Volcanoplot 
mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color=color)) + 
  geom_point(size=1, alpha=0.5) + theme_classic() + 
  ggtitle("R / NR")+
  scale_color_manual(values = c("red","blue",'grey')) +
  geom_vline(xintercept = c(-log2(1.2), log2(1.2)), color='red') +
  geom_hline(yintercept = -log10(0.05), color='blue') + theme_classic() 


```



```{r, eval=FALSE}
## R / NR in CTL   
obj_srt_CD8A_positive$response = factor(obj_srt_CD8A_positive$orig.ident)
levels(obj_srt_CD8A_positive$response) = c("NR","NR","R","NR","R","NR")

Idents(obj_srt_CD8A_positive) = "sample"
obj = subset(obj_srt_CD8A_positive, idents="CTL")

Idents(obj) = 'response'
logfc=log2(1)
mks =FindMarkers(obj, ident.1 = 'R', ident.2 = 'NR', 
                 logfc.threshold = logfc)
pval=0.05
fc=1.2
mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                               ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))

mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))

mks = mks %>% mutate(color = ifelse(is.na(p_val_adj), "no_sig",
                                    ifelse(log10(p_val_adj) <= log10(0.05) & 
                                             avg_log2FC >= log2(1.2), "UP", 
                                           ifelse(log10(p_val_adj) <= log10(0.05) & 
                                                    avg_log2FC <= -log2(1.2), "DN", "no_sig"))))
mks$color = factor(mks$color, levels = c("UP","DN","no_sig"))

# Volcanoplot 
mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color=color)) + 
  geom_point(size=1, alpha=0.5) + theme_classic() + 
  ggtitle("R / NR")+
  scale_color_manual(values = c("red","blue",'grey')) +
  geom_vline(xintercept = c(-log2(1.2), log2(1.2)), color='red') +
  geom_hline(yintercept = -log10(0.05), color='blue') + theme_classic() 

mks.ctl = mks
```


```{r, eval=FALSE}
## R / NR in TAK   
obj_srt_CD8A_positive$response = factor(obj_srt_CD8A_positive$orig.ident)
levels(obj_srt_CD8A_positive$response) = c("NR","NR","R","NR","R","NR")

Idents(obj_srt_CD8A_positive) = "sample"
obj = subset(obj_srt_CD8A_positive, idents="TAK")

Idents(obj) = 'response'
logfc=log2(1)
mks =FindMarkers(obj, ident.1 = 'R', ident.2 = 'NR', 
                 logfc.threshold = logfc)
pval=0.05
fc=1.2
mks = mks %>% mutate(DE=ifelse(avg_log2FC >= log2(fc) & p_val_adj < pval, 'UP',
                               ifelse(avg_log2FC <= -log2(fc) & p_val_adj < pval, 'DN','no_sig')))

mks$DE = factor(mks$DE, levels = c('UP','DN','no_sig'))

mks = mks %>% mutate(color = ifelse(is.na(p_val_adj), "no_sig",
                                    ifelse(log10(p_val_adj) <= log10(0.05) & 
                                             avg_log2FC >= log2(1.2), "UP", 
                                           ifelse(log10(p_val_adj) <= log10(0.05) & 
                                                    avg_log2FC <= -log2(1.2), "DN", "no_sig"))))
mks$color = factor(mks$color, levels = c("UP","DN","no_sig"))
mks.tak = mks
# Volcanoplot 
mks %>% ggplot(aes(avg_log2FC, -log10(p_val_adj), color=color)) + 
  geom_point(size=1, alpha=0.5) + theme_classic() + 
  ggtitle("R / NR")+
  scale_color_manual(values = c("red","blue",'grey')) +
  geom_vline(xintercept = c(-log2(1.2), log2(1.2)), color='red') +
  geom_hline(yintercept = -log10(0.05), color='blue') + theme_classic() 


```


