---
title: "Takeda PD1 + NK set merged"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: F
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)
```

# Merging condition  

4 from Elena's set    
MT30273.raw.23.05.16.rds    
MT30271.raw.23.05.16.rds    
MT30269.raw.23.05.16.rds    
MT30212.raw.23.05.16.rds    
    
2 from Patrick's set    
Takeda.NK.30298.23.08.17.rds    
Takeda.NK.30293.23.08.17.rds    
    
    


```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>

# Data Overview     

Lymphoid cells from CTL and TAK samples collected from 6 patients dataset   

Filtering  
- doublet removal   
- UMI : 500 < & < 20,000    
- mt content <= 25    
- ribosomal gene regression out    


# Lymphoid Data {.tabset}    


```{r}
dir <- "~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
obj.srt = readRDS(paste0(dir,"rds/obj.cond6_merged.lymphoid.rds"))
```


```{r}
obj.srt@meta.data[,c("orig.ident","sample")] %>% table() %>% data.frame() %>% tidyr::spread(sample, Freq) %>% DT::datatable()
```

<br>

## UMAP   
```{r}
DimPlot(obj.srt, group.by = "orig.ident")
```

## UMAP in split mode    
```{r, fig.width=12, fig.height=4}
DimPlot(obj.srt, group.by = "orig.ident", split.by = "orig.ident")
```



```{r, eval=FALSE}
## Gene expression  

genes = c("CD3E","CD8A","CD4","CD79A", "NKG7","TCF7","PTPRC")
FeaturePlot(obj.srt, features = genes, ncol = 3, slot = "data", pt.size = 0.1)
```
<br><br>

# Integrated data       

```{r, eval=F}
obj.list <- SplitObject(obj.srt, split.by = "orig.ident")

# normalize and identify variable features for each dataset independently
ifnb.list <- lapply(X = obj.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = ifnb.list)

#features 
# Perform integration (two steps)
obj.anchors <- FindIntegrationAnchors(object.list = ifnb.list, 
                                      anchor.features = features)
# this command creates an 'integrated' data assay
obj.intgr <- IntegrateData(anchorset = obj.anchors)

# Perform an integrated analysis

# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(obj.intgr) <- "integrated"

# Run the standard workflow for visualization and clustering
obj.intgr <- ScaleData(obj.intgr, verbose = FALSE)
obj.intgr <- RunPCA(obj.intgr, npcs = 30, verbose = FALSE)
obj.intgr <- RunUMAP(obj.intgr, reduction = "pca", dims = 1:30)
obj.intgr <- FindNeighbors(obj.intgr, reduction = "pca", dims = 1:30)
obj.intgr <- FindClusters(obj.intgr, resolution = c(0.1,0.2,0.5,0.8))

obj.intgr %>% saveRDS(paste0(dir,"rds/obj.cond6_merged.lymphoid_intgr.rds"))
```


```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
obj.intgr = readRDS(paste0(dir,"rds/obj.cond6_merged.lymphoid_intgr.rds"))
```


## UMAP   
```{r, fig.height=7, fig.width=8, out.width="60%"}
obj.intgr %>% DimPlot(group.by = 'orig.ident')
```

```{r}
obj.intgr %>% DimPlot(group.by = 'orig.ident', split.by = "orig.ident", ncol = 3, pt.size = 0.1)
```

## Clustering of resolution 0.1    
```{r, fig.height=7, fig.width=8, out.width="60%"}
obj.intgr %>% DimPlot(group.by = 'integrated_snn_res.0.1', pt.size = 0.1)
```


```{r, fig.height=7, fig.width=8, out.width="60%"}
obj.intgr %>% DimPlot(group.by = 'integrated_snn_res.0.1', pt.size = 0.1, label = T)
```

## Marker genes for resolution 0.1   

```{r, eval=FALSE}
find_and_save_markers <- function(obj.srt, cluster_id, logfc_threshold = 1.2, 
                                  test_method = 'wilcox', min_percent = 0.25) {
  Idents(obj.srt) = cluster_id
  all.markers = FindAllMarkers(obj.srt, logfc.threshold = log2(logfc_threshold), 
                               only.pos = TRUE, 
                               test.use = test_method, min.pct = min_percent)
  all.markers$cluster_fullname = paste0("cluster",all.markers$cluster)
  return(all.markers)
}
```

```{r, eval=FALSE}
intgr.mks = find_and_save_markers(obj.srt = obj.intgr, cluster_id = "integrated_snn_res.0.1")
intgr.mks %>% write.csv(paste0(dir,"data/mks/","lymphoid_intgr_res.0.1.mks.csv"))
```

```{r}
intgr.mks =read.csv(paste0(dir,"data/mks/","lymphoid_intgr_res.0.1.mks.csv"), row.names = 1)
```

```{r}
intgr.mks %>% DT::datatable(extensions = "Buttons", 
                options = list(autoWidth = TRUE, 
                               fixedHeader = TRUE,
                               dom="Bfrtip", buttons=c("csv","excel")))

```

## Top 10 marker genes heatmap   
```{r}
genes =intgr.mks %>% group_by(cluster_fullname) %>%  top_n(10, avg_log2FC) %>% select(gene) %>% pull()
obj.intgr@meta.data$RNA_snn_res.0.1 = factor(obj.intgr@meta.data$RNA_snn_res.0.1)
levels(obj.intgr@meta.data$RNA_snn_res.0.1) = c(0:9)
Idents(obj.intgr) = "RNA_snn_res.0.1"
DoHeatmap(obj.intgr, features = genes) +theme(text = element_text(size = 5))
```


## Top 5 marker genes dotplot   

```{r, fig.width=8, fig.height=12}
genes =intgr.mks %>% group_by(cluster_fullname) %>%  top_n(5, avg_log2FC) %>% select(gene) %>% pull()
obj.intgr@meta.data$RNA_snn_res.0.1 = factor(obj.intgr@meta.data$RNA_snn_res.0.1)
levels(obj.intgr@meta.data$RNA_snn_res.0.1) = c(0:9)
Idents(obj.intgr) = "RNA_snn_res.0.1"
DotPlot(obj.intgr, features = rev(genes), cols = c("grey","red")) +coord_flip()
```

## Suggested T cell types for clsuter    
```{r}
Idents(obj.intgr) = "RNA_snn_res.0.1"
genes = c("CD3D","CD8A","CD4")
#genes = c("CD3D","CD8A","CD4","NKG7","GZMA","GZMB","GZMH")
VlnPlot(obj.intgr, features = genes, stack = T, flip = T)
```

```{r}
DotPlot(obj.intgr, features = rev(genes), cols = c('grey','red')) + coord_flip()
```




**Suggested T cell category**   

| cluster  | CD3D  | CD8A  | CD4  | T category |
|----------|-------|-------|------|------------|
| cluster0 |   +   |   +   |   +  | CD3D+CD8+CD4+  |
| cluster1 |   +   |   +   |   -  | CD3D+CD8+  |
| cluster2 |   +   |   -   |   -  |   CD3D+    |
| cluster3 |   -   |   -   |   -  |   not T    |
| cluster4 |   -   |   -   |   -  |   not T    |
| cluster5 |   -   |   -   |   -  |   not T    |
| cluster6 |   +   |   -   |   +  | CD3D+CD4   |
| cluster7 |   -   |   -   |   -  |   not T    |
| cluster8 |   +   |   +   |   -  | CD3D+CD8+  |
| cluster9 |   -   |   -   |   -  |   not T    |


## CD8+ and CD4+ population   

```{r}
genes=c("CD8A","CD4")
FeaturePlot(obj.intgr, features = genes, pt.size = 0.2)
```
<br><br>

# GZMA level in CD8+T cells    

CD8A > 0 cells only   

```{r}
# obj.srt data
genes=c("CD8A","CD4","GZMA")
# genes.exp = data.frame(t(data.frame(obj.intgr@assays$integrated@data[genes,], check.names = F)))
genes.exp = data.frame(t(data.frame(obj.srt@assays$RNA@data[genes,], check.names = F)))
genes.exp.meta = cbind(genes.exp, obj.srt@meta.data[rownames(genes.exp),])

#genes.exp[1:3,]
# genes.exp.meta %>% 
#   filter(CD8A > 0) %>% 
#   ggplot(aes(orig.ident, GZMA, color = orig.ident)) +
#   geom_boxplot(outlier.size = 0.1) +
#   xlab("") +
#   theme_bw() +
#   theme(legend.title = element_blank(),
#         plot.title = element_text(hjust = 0.5, size = 16)) +  
#   ggtitle("GZMA level in CD8+T cells") 


genes.exp.meta %>% 
  filter(CD8A > 0) %>% 
  ggplot(aes(orig.ident, GZMA, color = orig.ident)) +
  geom_violin(draw_quantiles = T) +
  xlab("") +
  theme_bw() +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +  
  ggtitle("GZMA level in CD8+T cells") 

```

