---
title: "Takeda PD1 + NK set merged"
author: "Sung Rye Park"
date: "`r format(Sys.Date())`"
output:  
  rmdformats::robobook: 
    code_folding: hide 
    number_sections: TRUE
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, results = 'asis')
options(warn = F)

```

# Merging condition  

4 from Elena's set    
MT30273.raw.23.05.16.rds    
MT30271.raw.23.05.16.rds    
MT30269.raw.23.05.16.rds    
MT30212.raw.23.05.16.rds    
    
2 from Patrick's set    
Takeda.NK.30298.23.08.17.rds    
Takeda.NK.30293.23.08.17.rds    
    
    


```{r, echo=TRUE}
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
```
<br>
<br>


Selected samples (after filtering)    

| orig.ident | CTL  | TAK  |
|------------|------|------|
|   30293    | 3976 | 5182 |
|   30298    | 9620 | 6054 |
|  MT30212   | 522  | 463  |
|  MT30269   | 6185 | 5551 |
|  MT30271   | 5602 | 6412 |
|  MT30273   | 3031 | 1569 |


Filtering  
- doublet removal   
- UMI : 500 < & < 20,000    
- mt content <= 25    
- ribosomal gene regression out    


# Data   
```{r}
dir='~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/'
obj.srt =readRDS(paste0(dir,'rds/Takeda_6conditions.23.12.01.rds'))
```


Total number of cells: 54167    


```{r,eval=FALSE}
# Total number of cells 
obj.srt@meta.data %>% nrow()
```


# Heatmaps    


## 0.2  

```{r}
dir="~/Desktop/DF/DFCI_Paweletz/2023_Takeda_PD1_NK/"
```


```{r}
resolution_number = 0.2
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=12, fig.height=12}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 0.4  

```{r}
resolution_number = 0.4
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=14, fig.height=14}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 0.8  

```{r}
resolution_number = 0.8
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=16, fig.height=20}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 8))
```


## 1  


```{r}
resolution_number = 1
res= paste0("RNA_snn_res.", resolution_number)
mks = read.csv(paste0(dir,"data/mks/Merge6_", res, ".markers.csv"), row.names = 1)
genes = mks %>% group_by(cluster) %>% top_n(5, avg_log2FC) %>% select(gene) %>% pull()
```

```{r, fig.width=22, fig.height=28}
Idents(obj.srt) = res
DoHeatmap(obj.srt, features = genes) +theme(text = element_text(size = 10))
```